---
title: "Further quality control"
author: "Karsten Bach"
date: '`r Sys.Date()`'
output:
  html_notebook:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: journal
    highlight: tango
    code_folding: hide
---
***

In this document I will further clean up the data and remove cells that most likely represent doublets, damaged cells or red blood cell contamination.


# Removal of doublets
The following strategy for doubled removal was employed, mainly following an approach from __Pijuana-Sala, Griffiths and Guibentif et al.__

1. Estimate probability of cell being a doublet per sample using doubletCells from scran.
2. Compute per sample clusters based on all genes based on a SNNGraph (k=10) and walktrap for community detection (steps=2).
3. Clusters with high average doublet probability are identified as doublets.
4. All cells and samples (incl. doublets) are integrated using fastMNN, cells that cluster with "known" doublets cells are guilty by association.

## Per sample clusters
- Below are the per sample clusters as well as the doublet scores per cell.
- WKBR76.5e3 has clear outliers which I suspect to be damaged cells, see end of the document.
- WKBR75.4bMG3 has clearly been overloaded, however there is structure which is not as obvious in the small plot. Will need to double check after doublet removal how this looks.
```{r, message=FALSE,fig.height=17,fig.width=17}
# Identifying doublets with doubletCluster
library(ggplot2)
library(scran)
library(cowplot)
library(wesanderson)
library(viridis)
theme_set(theme_cowplot())

# Read in data
out <- readRDS("../data/combined_Robjects/DoubletData.rds")

# Visualizing the clustering
cluster_plots <- lapply(names(out), function(nm) {
		   x <- out[[nm]]
		   p <- ggplot(x, aes(x=SubUMAP1, y=SubUMAP2, color=factor(Cluster))) +
			       geom_point(size=0.7) +
			       scale_color_manual(values=wes_palette("FantasticFox1",length(unique(x$Cluster)),type="continuous")) +
			       theme(legend.position="none", panel.background = element_rect(colour = "black", linetype = 1, size = 0.5),
				     axis.line = element_blank(),
				     axis.ticks = element_blank(),
				     axis.text = element_blank(),
				     axis.title = element_blank(),) +
			       ggtitle(paste0(nm," (n=",nrow(x),", k=",length(unique(x$Cluster)),")")) 
			   return(p)
})
names(cluster_plots) <- names(out)
plot_grid(plotlist=cluster_plots)

# Visualizing the doublet scores
score_plots <- lapply(names(out), function(nm) {
		   x <- out[[nm]]
		   p <- ggplot(x, aes(x=SubUMAP1, y=SubUMAP2, color=log2(DbltScore+1))) +
			       geom_point(size=0.7) +
			       scale_color_viridis() +
			       theme(legend.position="none", panel.background = element_rect(colour = "black", linetype = 1, size = 0.5),
				     axis.line = element_blank(),
				     axis.ticks = element_blank(),
				     axis.text = element_blank(),
				     axis.title = element_blank(),) +
			       ggtitle(paste0(nm," (n=",nrow(x),", k=",length(unique(x$Cluster)),")")) 
			   return(p)
})
names(score_plots) <- names(out)
plot_grid(plotlist=score_plots)
```

## Identify clusters per samples that represent doublets
- Clusters that had a higher than median + 5xmad doublet score where called as doublet cells
```{r, message=FALSE, fig.width=10}
# Identifying doublets based on doubletCell
# Add sample index to data frames
outmod <- lapply(names(out), function(x) {
		dfs <- out[[x]]
		dfs$Sample <- x
		return(dfs)
})

outdf <- do.call(rbind,outmod)

# Compute median scores per cluster and sample
medscores <- aggregate(outdf$DbltScore, list(outdf$Sample, outdf$Cluster), median)
names(medscores) = c("sample", "cluster", "median.score")

medscores$n.cells = sapply(1:nrow(medscores), function(row){
			       sum(outdf$Cluster == medscores$cluster[row] & outdf$Sample == medscores$sample[row])
})

medscores$fracCells = sapply(1:nrow(medscores), function(row){
				  medscores$n.cells[row]/sum(medscores$n.cells[medscores$sample == medscores$sample[row]])
})


# Define doublets based on fraction and median doublet score
medscores$outlier <- scater::isOutlier(medscores$median.score,nmad=5,type="higher")
medscores$isDoublet <- medscores$outlier & medscores$fracCells <= 0.05

# Plot from Johnny with the significant clusters highlighted
ggplot(medscores, aes(x = sample, y = log2(median.score + 1), col = factor(cluster))) +
  geom_point(data = medscores[medscores$isDoublet,]) +
  geom_jitter(data = medscores[!medscores$isDoublet,], col = "darkgrey", size = 0.4, width = 0.2, height = 0) +
  guides(color="none") +
  labs(x = "Sample", y = "log2(cluster median score + 1)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

medscores.dbls <- medscores[medscores$isDoublet,]
smp.dbltCluster <- paste(medscores.dbls$sample,medscores.dbls$cluster,sep="_")
```


## Visualizing the doublet called clusters
- most clusters show the typical tear shape in the UMAP
- however we can already see here that there might be more doublet clusters than we identified above
```{r, message=FALSE, fig.height=13, fig.width=13}
smps <- names(out)[names(out) %in% medscores.dbls$sample]
doublet_plots <- lapply(smps, function(nm) {
		   x <- out[[nm]]
		   p <- ggplot(x, aes(x=SubUMAP1, y=SubUMAP2)) +
			       geom_point(color="grey80",size=0.7) +
			       geom_point(data=x[x$Cluster %in% medscores.dbls[medscores.dbls$sample==nm,"cluster"],], aes(color=factor(Cluster))) +
			       #                                scale_color_manual(values=wes_palette("FantasticFox1",length(unique(x$Cluster)),type="continuous")) +
			       theme(legend.position="none", panel.background = element_rect(colour = "black", linetype = 1, size = 0.5),
				     axis.line = element_blank(),
				     axis.ticks = element_blank(),
				     axis.text = element_blank(),
				     axis.title = element_blank(),) +
			       ggtitle(paste0(nm," (n=",nrow(x),", k=",length(unique(x$Cluster)),")")) 
			   return(p)
})
names(doublet_plots) <- smps
plot_grid(plotlist=doublet_plots)
```

## Doublet clusters have a larger library size
```{r, message=FALSE}
pD <- readRDS("../data/combined_Robjects/ExpressionList_QC_norm.rds")[["phenoData"]]
library(dplyr)
outdf <- left_join(outdf,pD[,c("barcode","UmiSums","GenesDetected","prcntMito")])
outdf$sampClust <- paste0(outdf$Sample,"_",outdf$Cluster)
outdf$isDoublet <- ifelse(outdf$sampClust %in% smp.dbltCluster,"Doublet","Singlet")
ggplot(outdf, aes(x=isDoublet,y=log10(UmiSums))) +
       geom_boxplot() 
```

## Cluster batch corrected data
- Next we will identify cells that are guilty-by-association by clustering the full dataset including the above identified doublet cells and find clusters with a high frequency of doublets
```{r, message=FALSE, warning=FALSE, cache=TRUE}
# Read in BC PCA
m.cor <- readRDS("../data/combined_Robjects/CorrectedPCA_preDoublet.rds")
igr <- buildSNNGraph(m.cor,d=NA, transposed=TRUE)
# Overcluster
cl <- igraph::cluster_walktrap(igr,steps=2)
cluster <- cl$membership
clusDf <- data.frame("barcode"=rownames(m.cor),
		     "fullCluster"=cluster)
outdf <- dplyr::left_join(outdf,clusDf)
```


## Guilty-by-association clusters
- All clusters with more than median+ 2x mad fraction of doublets are additionally identifed as doublet clusters
```{r, message=FALSE, warning=FALSE}
smry <- group_by(outdf, fullCluster, isDoublet) %>%
    summarize(NType=n()) %>%
    mutate(Prop=NType/sum(NType))

doubDef <- scater::isOutlier(smry$Prop[smry$isDoublet=="Doublet"],nmads=2,type="higher")
dblClusters <- smry$fullCluster[smry$isDoublet=="Doublet"][doubDef]
thrsld <- attributes(doubDef)[[1]][2]

ggplot(smry, aes(x=fullCluster, y=Prop, fill=isDoublet)) +
    geom_bar(stat="identity") +
    geom_hline(yintercept=1-thrsld) +
    scale_fill_manual(values=c("black","red"))

ump <- read.csv("../data/combined_Robjects/UMAP_corrected_preDoublet.csv",row.names=1)
outdf <- left_join(outdf,ump)

ggplot(outdf, aes(x=UMAP1, y=UMAP2, color=outdf$fullCluster %in% dblClusters)) +
    geom_point() +
    scale_color_manual(values=c("black","red")) +
    theme(legend.position="none")
```

## Highlighting all cells identified as doublets
```{r, message=FALSE}
dbl.barcodes <- outdf$barcode[outdf$isDoublet=="Doublet" | outdf$fullCluster %in% dblClusters]

outdf$isDoubletFinal <- outdf$barcode %in% dbl.barcodes
ggplot(outdf, aes(x=UMAP1, y=UMAP2, color=outdf$isDoubletFinal)) +
    geom_point() +
    scale_color_manual(values=c("black","red")) +
    theme(legend.position="none")
```

- As expected sample with more cells tend to have a higher frequency of doublets.
```{r, message=FALSE}
smry <- group_by(outdf, Sample, isDoubletFinal) %>%
    summarize(n=n()) %>%
    mutate(DoubletFraction=n/sum(n),totCells=sum(n)) %>%
    ungroup() %>%
    filter(isDoubletFinal)

ggplot(smry, aes(x=totCells,y=DoubletFraction)) +
    geom_point()
```

# Other outliers
- The first sample appears to have a cluster that is very distinct from everything else, these might represent damaged cells
- To try to identify damaged cells I looked at % of mitochondrial reads versus library size

```{r, message=FALSE, warning=FALSE}
m <- readRDS("../data/combined_Robjects/ExpressionList_QC_norm.rds")[["counts"]]
hbb.exp <- colSums(m[c("Hbb.bt","Hbb.bs"),])
hbb.exp <- data.frame("barcode"=names(hbb.exp),
		      "hbbExp"=hbb.exp)
outdf <- left_join(outdf,hbb.exp)
smry <- group_by(outdf, sampClust) %>%
    summarize(medUmi=median(UmiSums),
	      medprcntMito=median(prcntMito),
	      medGd=median(GenesDetected),
	      medHb=median(hbbExp)) %>%
    ungroup() %>%
    as.data.frame()

ggplot(smry, aes(x=medUmi,y=medprcntMito)) +
    geom_point() +
    scale_x_log10() +
    scale_y_log10() +
    geom_hline(yintercept=0.005) +
    xlab("Library Size") +
    ylab("Fraction of Mitochondrial Reads")
```

- However I couldn't identify a clear cutoff
- And also in the UMAPs below I am not fully convinced that they are all truly damaged cells
- I did label cells as "potentially damaged" and will keep that information for further analyses

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
cls <- smry[smry$medprcntMito<0.005,"sampClust"]
smps.dmgd <- unlist(lapply(strsplit(cls,split="_"),function(x) x[1]))
cls.dmgd <- unlist(lapply(strsplit(cls,split="_"),function(x) x[2]))

dmgd_plots <- lapply(1:length(smps.dmgd), function(i) {
		   nm <- smps.dmgd[i]
		   x <- out[[nm]]
		   p <- ggplot(x, aes(x=SubUMAP1, y=SubUMAP2)) +
			       geom_point(color="grey80",size=0.7) +
			       geom_point(data=x[x$Cluster %in% cls.dmgd[i],], aes(color=factor(Cluster))) +
			       #                                scale_color_manual(values=wes_palette("FantasticFox1",length(unique(x$Cluster)),type="continuous")) +
			       theme(legend.position="none", panel.background = element_rect(colour = "black", linetype = 1, size = 0.5),
				     axis.line = element_blank(),
				     axis.ticks = element_blank(),
				     axis.text = element_blank(),
				     axis.title = element_blank(),) +
			       ggtitle(paste0(nm," (n=",nrow(x),", k=",length(unique(x$Cluster)),")")) 
			   return(p)
})
names(dmgd_plots) <- smps.dmgd
plot_grid(plotlist=dmgd_plots)
dmgd <- outdf$barcode[outdf$sampClust %in% cls]
outdf$isPotDamaged <- outdf$barcode %in% dmgd
```

- Further I exclude clusters that show expression of hemoglobin
```{r, message=FALSE, warning=FALSE}
library(ggrepel)
ggplot(smry, aes(x=sampClust,y=medHb)) +
    geom_point() +
    xlab("Per-Sample Clusters") +
    ylab("Median Hbb Expression") +
    geom_label_repel(data=smry[smry$medHb>0.01,],aes(label=sampClust)) +
    theme(axis.text.x=element_blank())
rbcs <- smry[smry$medHb>0.01,"sampClust"]
rbcs <- outdf$barcode[outdf$sampClust %in% rbcs]
```
# Save
```{r, message=FALSE, warning=FALSE}
outdf$isRbc <- outdf$barcode %in% rbcs
qcMets <- outdf[,c("barcode","DbltScore","isRbc","isDoubletFinal","isPotDamaged")]
write.csv(qcMets,"../data/combined_Robjects/QC_Part2.csv")
```
