---
title: "Tumorigenesis"
author: "Karsten Bach"
date: '`r Sys.Date()`'
output:
  html_notebook:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: journal
    highlight: tango
    code_folding: hide
---
***

# How do we define tumorigenesis

We are interested in finding patterns that are associated with tumorigenesis.
For this we need to have a (per-sample) covariate that represents the process of tumorigenesis, however, the only one we have is the age of the animal.
Despite being highly correlated with the onset of tumorigenesis, the variability in latency would confound the analysis.
Hence, we want to estimate some sort of "biological time", ideally directly from the data.

```{r, message=FALSE, warning=FALSE}
library(scran)
library(scater)
library(dplyr)
library(knitr)
library(ggplot2)
library(Rtsne)
library(cowplot)
library(Matrix)
library(viridis)
library(knitr)
library(wesanderson)
library(DT)
library(edgeR)
source("../functions.R")
theme_set(theme_cowplot())

# Load Data
sce.full <- readRDS("../../data/Integrated/Robjects/SCE_combined_celltypes.rds")
sce <- sce.full[,sce.full$Experiment=="Tumorigenesis"]
# Excluding WT and p53 wt
sce <- sce[,!sce$Condition %in% c("CBLT","CBLA1","WKBR76.5f","WKBR76.5c")]
sce$CellTypes <- sce$CellTypesFinal
sce$Batch <- as.character(sce$Batch)
sce$Age <- as.character(sce$Age)
colnames(sce) <- sce$barcode ### I have to fix this somewhere else
pD <- data.frame(colData(sce))
set.seed(42)
pD <- pD[sample(nrow(pD),nrow(pD)),]
```

## Sample Overview
These are the ages at which the WKBR animals were collected.
This includes two tumors in the last two groups of two.
```{r, message=FALSE, fig.width=6, fig.height=3}
sumry <- group_by(pD, Batch, Age, Condition) %>%
    summarize(n=n())

sumry <- sumry[grepl("WKBR",sumry$Condition),]
timeplt <- ggplot(sumry, aes(x=as.numeric(Age), y=0)) +
    geom_segment(aes(x=25, y=0, xend=60, yend=0,
		     color=NULL),
                  arrow=arrow(length=unit(0.5, "cm"))) +
    geom_segment(aes(x=30, y=0.05, xend=30, yend=0,
		     color=NULL),
                  arrow=arrow(length=unit(0.3, "cm"))) +
    geom_dotplot(aes(fill=Batch, group=Batch), binwidth=1, 
		 stackgroups=TRUE, binpositions="all") +
    xlim(25,60) +
    ylim(0,.5) +
    scale_fill_manual(values=c("grey20","grey60","black"))  +
    xlab("Age") +
    ggtitle("Timepoints of Collection") +
    theme(axis.title.y=element_blank(),
	  axis.line.x=element_blank(),
	  axis.line.y=element_blank(),
	  axis.text.y=element_blank(),
	  legend.position="none",
	  axis.ticks.y=element_blank(),
	  axis.ticks.x=element_line(size=1, linetype="solid"),
	  axis.ticks.length.x=unit(.5,"cm"))
timeplt
```

# Ideas
We have mainly discussed two broad ideas.

#### Based on Expression

- In this approach we use the expression data directly in a way to derive some per-sample estimate of how close to tumor formation a sample is
- Advantages
    + Based on fairly light assumptions
    + Simple
    + Easy to interpret
    + Depending on which method we use you could predict the "tumor stage" of external samples
- Disadvantages
    + Potential circularity when interpreting the results

#### Based on Genomic Alterations

- In this approach we first estimate CNAs based on the gene expression and then create a summary statistic that summarizes how "cancerous" the cells already are 
- Advantages
    + The "tumor time covariate" will be more independent from the data that will actually be analysed downstream
- Disadvantages
    + Not quite clear how to get a meaningful (in terms of "cancerousness") summary of CNA profiles across multiple cells
    + Relationship between cancerousness and CNAs not quite clear
    + Estimation of CNA from 10x data not trivial

# Expression based
In this section I will explore approaches that directly utilize the gene expression to estimate a pseudo-tumor time.

## PCAs {.tabset .tabset-fade}
After having compared PCAs on the various compartments, I found the one on either the whole data or epithelium only the most compelling ones.
The disadvantage of using a subset is that it might appear biased, however I think it's hard to argue against using all epithelial.
On the plus side using a subset might make analyses of other subsets independent of this ordering.
However, this also only true to a certain extent if we are ignoring potential cross-talk between the signal in the epithelial compartment and the other compartments.
```{r, message=FALSE, warning=FALSE}
getPC <- function(sce, getPCObject=FALSE) {
    sce.sum <- aggregateAcrossCells(sce,
			  id=DataFrame(Condition=sce$Condition),
			  use_exprs_values="counts",
			  average=FALSE)

    sce.counts <- counts(sce.sum)
    y.sce <- DGEList(counts=sce.counts, samples=data.frame(colData(sce.sum)))
    keep <- filterByExpr(y.sce)
    y.sce <- y.sce[keep,]
    y.sce <- calcNormFactors(y.sce)

    logcpm <- cpm(y.sce, log=TRUE)
    # 
    pD <- data.frame(colData(sce.sum))
    #
    pcs <- prcomp(t(logcpm))
    pca <- data.frame(pcs$x)
    pca[,"Condition"] <- pD$Condition
    pca$Age <- pD$Age
    pca$Batch <- pD$Batch
    # pca$CellTypes <- y$samples$CellTypes
    rownames(pca) <- pca$Condition
    if (getPCObject) {
	return(pcs)
    } else {
	return(pca)
    }
}

plotPC <- function(pca) {
    pcs <- as.matrix(pca[,grepl("PC",colnames(pca))])
    vars <- apply(pcs,2,var)
    rel.vars <- round(vars/sum(vars)*100,1)
    # 
    library(ggrepel)
    pout <- ggplot(pca, aes(x=PC1, y=PC2, color=Age,label=Condition,shape=Batch)) +
	    geom_text_repel(size=3) +
	    geom_point(size=2) +
	    xlab(paste0("PC1(",rel.vars[1],"%)")) +
	    ylab(paste0("PC2(",rel.vars[2],"%)")) +
	    theme_bw() +
	    coord_fixed(1)
    return(pout)
}

```
# All Cells

```{r, message=FALSE, warning=FALSE}
pcs.all <- getPC(sce) 
p.all <- plotPC(pcs.all) + ggtitle("Full Data")
```


## Ordering based on PC1
We can then create the same plot as before but this time with PC1 as being the X-Axis.
```{r, message=FALSE, warning=FALSE, fig.width=7, fig.height=3.5}
sumry <- arrange(sumry,Age)
sumry$Condition <- factor(sumry$Condition,
			  levels=sumry$Condition)
cols <- data.frame(Age=unique(sumry$Age))
cols$cols <- plasma(nrow(cols),begin=1,end=0)

sumry <- left_join(sumry,cols)

timeplt_byCond <- ggplot(sumry, aes(x=as.numeric(Age), y=0)) +
    geom_segment(aes(x=25, y=0, xend=60, yend=0,
		     color=NULL),
		 arrow=arrow(length=unit(0.5, "cm"))) +
geom_segment(aes(x=30, y=0.05, xend=30, yend=0,
		 color=NULL),
	     arrow=arrow(length=unit(0.3, "cm"))) +
geom_dotplot(aes(fill=Condition, group=Condition), binwidth=1, 
	     stackgroups=TRUE, binpositions="all") +
xlim(25,60) +
ylim(0,.5) +
xlab("Age") +
scale_fill_manual(values=sumry$cols) +
ggtitle("Timepoints of Collection") +
theme(axis.title.y=element_blank(),
      axis.line.x=element_blank(),
      axis.line.y=element_blank(),
      axis.text.y=element_blank(),
      legend.position="none",
      axis.ticks.y=element_blank(),
      axis.ticks.x=element_line(size=1, linetype="solid"),
      axis.ticks.length.x=unit(.5,"cm"))

pcs.all$PC1 <- -pcs.all$PC1
pcs.all$ptime <- (pcs.all$PC1-min(pcs.all$PC1))/max(pcs.all$PC1-min(pcs.all$PC1)) * 100

sumry <- left_join(sumry,pcs.all[,c("Condition","ptime")])
sumry$Condition <- factor(sumry$Condition,
			  levels=sumry$Condition)

pseudotimeplt_byCond <- ggplot(sumry, aes(x=ptime, y=0)) +
    geom_segment(aes(x=0, y=0, xend=110, yend=0,
			 color=NULL),
		     arrow=arrow(length=unit(0.5, "cm"))) +
    geom_dotplot(aes(fill=Condition, group=Condition), binwidth=1, 
		 stackgroups=TRUE, binpositions="all", dotsize=2) +
    xlim(0,110) +
    ylim(0,.5) +
    xlab("Pseudotime") +
    scale_fill_manual(values=sumry$cols) +
    ggtitle("Inferred Order") +
    theme(axis.title.y=element_blank(),
	  axis.line.x=element_blank(),
	  axis.line.y=element_blank(),
	  axis.text.y=element_blank(),
	  legend.position="none",
	  axis.ticks.y=element_blank(),
	  axis.ticks.x=element_line(size=1, linetype="solid"),
	  axis.ticks.length.x=unit(.5,"cm"))

timeplt_byCond
pseudotimeplt_byCond
```

# Sanity Checks

There are certain things we can assume should be true if PC1 corresponds to tumor development.

1. It should be positively correlated with age
2. Similarity to wildtype data should decrease along PC1
3. The MPA-DMBA timeseries data should be mapped according to time
4. "Meaningful" loadings in PC1

## Correlation with Age
Animals with higher age on average have a higher PC1 values.
```{r, message=FALSE, warning=FALSE}
cr <- cor.test(as.numeric(pcs.all$Age), pcs.all$PC1)

txt <- data.frame("PC1"=c(-40,-35),
	   "Age"=c(48,47),
	   "Lbls"=c(sprintf("rho == %0.2f",cr$estimate),
		    sprintf("PVal == %0.2f", cr$p.value)))
ggplot(pcs.all, aes(x=PC1, y=as.numeric(Age))) +
    geom_point(size=3,fill="grey50",color="black",pch=21) + 
    geom_text(data=txt, aes(label=Lbls),parse=TRUE,size=6) +
    xlab("-PC1") +
    ylab("Age [w]") +
    theme_pub()
    #     geom_text(x=-50,y=47.5,label=paste0("Cor: ",cr))
```

## MPA-DMBA Data

I here compare the Brca1 timeline experiment to the DMBA timeline.
As a reminder in the DMBA timeline we have collected several timepoints during tumor formation after MPA-DMBA treatment.
The experiment contained untreated (T0), MPA-only (T1) and then animals in two weeks interval up to 10 weeks post last MPA.
It also contained three tumor in total from group T5 and T6.

```{r, message=FALSE, warning=FALSE}
dmba.data <- readRDS("../../data/Tumorigenesis/Robjects/DMBA_Bulk.rds")
pDat <- dmba.data[["pDat"]]
dmba.counts <- dmba.data[["counts"]]
kable(table(pDat$Timepoint,pDat$Tumor))
```

Below I recompute the PCA on the single-cell data on all genes that were also measured in the DMBA data (this only excludes 70 genes).
I then apply the same rotation from the PCA to the DMBA bulk RNA-sequencing data and see where it maps to along PC1.
I have visualised this below both as PCA plot as well as along the timeline.
```{r, message=FALSE, warning=FALSE}
# ReDo the PCA on all samples # and switch to ENSMBL IDs
sce.sum <- aggregateAcrossCells(sce,
		  id=DataFrame(Condition=sce$Condition),
		  use_exprs_values=c("counts"),
		  average=FALSE)
rD <- data.frame(rowData(readRDS("../../data/Tumorigenesis/Robjects/SCE.rds")))
rD <- rD[rownames(sce.sum),]
rownames(sce.sum) <- rD$ID
#Remove genes that are not in the DMBA data
rmgenes <- setdiff(rownames(sce.sum),rownames(dmba.counts)) # 71
sce.sum <- sce.sum[!rownames(sce.sum) %in% rmgenes,]
sce.counts <- counts(sce.sum)
y.sce <- DGEList(counts=sce.counts, samples=data.frame(colData(sce.sum)))
keep <- filterByExpr(y.sce)
y.sce <- y.sce[keep,]
y.sce <- calcNormFactors(y.sce)

logcpm <- cpm(y.sce, log=TRUE)
# 
pcs <- prcomp(t(logcpm))
```

```{r, message=FALSE, warning=FALSE}
# 1. Subset bulk RNA data to same genes
dmba.counts <- dmba.counts[rownames(logcpm),]
# Set up DGEList
y.dmba <- DGEList(counts=dmba.counts,samples=pDat)
# y <- y[,!y$samples$Tumor]
# Normalize
y.dmba <- calcNormFactors(y.dmba)
# Get Log-Cpm
dmba.logcpm <- cpm(y.dmba,log=TRUE)
# Predict into previous PC space
pcpredict <- predict(pcs,t(dmba.logcpm))
res <- y.dmba$samples[,c("Sample","Timepoint","Tumor")]
res$PC1 <- -pcpredict[,1]
res$PC2 <- pcpredict[,2]
colnames(res)[1] <- "Condition"
fplot <- data.frame("Condition"=sce.sum$Condition,
		    "Tumor"=grepl("TM",sce.sum$Condition),
		    "Timepoint"=sce.sum$Age,
		    "PC1"=-pcs$x[,"PC1"],
		    "PC2"=pcs$x[,"PC2"],
		    "Experiment"="WKBR")
res$Experiment <- "DMBA"


fplot <- rbind(fplot,res)

minpc1 <- min(fplot[fplot$Experiment=="WKBR","PC1"])
maxpc1 <- max(fplot[fplot$Experiment=="WKBR","PC1"]-minpc1)
fplot$ptime <- ((fplot$PC1-minpc1)/maxpc1) * 100

fplot.wkbr <- fplot[fplot$Experiment=="WKBR",]
fplot.dmba <- fplot[fplot$Experiment=="DMBA",]
# fplot.dmba <- left_join(fplot.dmba,res[,c("Condition","Tumor")])

cols <- data.frame("Timepoint"=unique(fplot.dmba$Timepoint))
cols$col <- plasma(nrow(cols), begin=1, end=0)
fplot.dmba <- left_join(fplot.dmba,cols)
fplot.dmba$col[fplot.dmba$Tumor] <- "Red"
fplot.dmba$Condition <- factor(fplot.dmba$Condition,
			       levels=fplot.dmba$Condition)

ggplot(fplot.dmba, aes(x=PC1, y=PC2)) +
     geom_point(data=fplot.wkbr, color="grey80") +
     geom_point(aes(color=Condition),size=2) +
     scale_color_manual(values=fplot.dmba$col) +
     theme_pub()



plt <- ggplot(fplot.dmba, aes(x=ptime, y=0)) +
    geom_segment(aes(x=0, y=0, xend=110, yend=0,
		     color=NULL),
		 arrow=arrow(length=unit(0.5, "cm"))) +
    geom_dotplot(data=fplot.wkbr,fill="grey80", binwidth=1, 
		 stackgroups=TRUE, binpositions="all", dotsize=1) +
    geom_dotplot(aes(fill=Condition), binwidth=1, 
		 stackgroups=TRUE, binpositions="all", dotsize=2) +
    xlim(0,110) +
    ylim(0,.5) +
    xlab("Pseudotime") +
    scale_fill_manual(values=fplot.dmba$col) +
    ggtitle("DMBA projected on Inferred Order") +
    theme(axis.title.y=element_blank(),
	  axis.line.x=element_blank(),
	  axis.line.y=element_blank(),
	  axis.text.y=element_blank(),
	  legend.position="none",
	  axis.ticks.y=element_blank(),
	  axis.ticks.x=element_line(size=1, linetype="solid"),
	  axis.ticks.length.x=unit(.5,"cm"))
plt

fplot.dmba$Timepoint <- as.character(fplot.dmba$Timepoint)
fplot.dmba$Timepoint[fplot.dmba$Tumor] <- "Tumor"
fplot.dmba$Timepoint <- factor(fplot.dmba$Timepoint,levels=c("T0","T1","T2",
							     "T3","T4","T5",
							     "T6","Tumor"))
ggplot(fplot.dmba, aes(x=Timepoint, y=PC1)) +
     geom_boxplot() +
     theme_pub()
```

## PC1 Loadings
Finally, we can directly look at the loadings of PC1 to see if the genes with high positive loadings can be related to the process of tumor formation.
In fact, we see cell division and innate immune response popping up.
The negative loadings are not so easy to interpret.
```{r, message=FALSE, warning=FALSE}
pcs <- getPC(sce, getPCObject=TRUE)


ld.pc1 <- pcs$rotation[,1]
ld.pc1.ord <- ld.pc1[order(ld.pc1,decreasing=FALSE)]
fp <- data.frame("Gene"=names(ld.pc1.ord),
		 "Loading"=ld.pc1.ord)
fp$Index <- c(1:nrow(fp))
library(ggrepel)

p.ld <- ggplot(fp, aes(x=Index, y=Loading)) +
    geom_point(color="grey80") +
    geom_point(data=fp[abs(fp$Loading)>0.02,],color="black") +
    geom_text_repel(data=fp[c(1:5,(nrow(fp)-5):nrow(fp)),], aes(label=Gene),size=5) +
    geom_hline(yintercept=0.02,color="red",lty="dashed") +
    geom_hline(yintercept=-0.02,color="red",lty="dashed") +
    ylab("PC1 Loading") +
    theme_pub() +
    coord_flip()

ps <- ld.pc1[ld.pc1 > 0.02]
ps <- ps[order(ps,decreasing=TRUE)]
neg <- ld.pc1[ld.pc1 < -0.02]
neg <- neg[order(neg)]

topLoads <- list("Positive Loading"=ps, "Negative Loading"=neg)# topGO
univrs <- names(ld.pc1)
out <- list()

for (i in seq_along(topLoads)) {
    gens <- names(topLoads[[i]])
    subst <- names(topLoads)[[i]]
    # Gene universe
    alG <- factor(as.numeric(univrs %in% gens))
    names(alG) <- univrs

    # ---- GOanalysis ----

    library(topGO)
    #prepare Data for topGO
    GO.data <- new("topGOdata", description="Lib GO",ontology="BP", allGenes=alG, 
		   annot=annFUN.org, mapping="org.Mm.eg.db",
		   nodeSize=20, ID="symbol")
    result.classic <- runTest(GO.data, statistic="fisher")
    res <- GenTable(GO.data, Fisher.classic=result.classic, orderBy="topgoFisher", topNodes=50, numChar=300)

    res$Fisher.classic <- as.numeric(res$Fisher.classic)
    res$Term <- factor(res$Term, levels=res$Term[order(res$Fisher.classic,decreasing=TRUE)])
    out[[subst]] <- res[1:15,] %>%  # this only works because they are ordered by fisher classic
			mutate(hitsPerc=Significant*100/length(gens)) %>% 
			ggplot(aes(y=hitsPerc, 
				   x=Term, 
				   fill=Fisher.classic
				   )) +
			    geom_col() +
			    theme_pub() +
			    expand_limits(x=0) +
			    labs(y="Hits (%)", x="GO term", colour="p value", size="Count") +
			    ylim(c(0,20)) +
			    scale_fill_viridis(begin=1,end=0) +
			    ggtitle(subst) +
			    coord_flip()
}

p.ld
```


```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
out[[2]]
out[[1]]
```

```{r, message=FALSE, warning=FALSE}
# Repeat PC -.-
pc.obj <- getPC(sce,getPCObject=TRUE) 
# Selecting p53 wt
sce.p53 <- sce.full[,sce.full$Condition %in% c("WKBR76.5f","WKBR76.5c")]
sce.sum <- aggregateAcrossCells(sce.p53,
		  id=DataFrame(Condition=sce.p53$Condition),
		  use_exprs_values=c("counts"),
		  average=FALSE)
sce.counts <- counts(sce.sum)
sce.counts <- sce.counts[rownames(pc.obj$rotation),]
y.sce <- DGEList(counts=sce.counts, samples=data.frame(colData(sce.sum)))
y.sce <- calcNormFactors(y.sce)
logcpm <- cpm(y.sce, log=TRUE)

pcpredict <- predict(pc.obj,t(logcpm))




pcs.all$PC1 <- pcs.all$PC1
pcs.all$SampleID <- rownames(pcs.all)
pcs.all$Tumor <- grepl("TM",pcs.all$SampleID)
pcs.all$Tumor <- ifelse(pcs.all$Tumor,"Tumor","Mammary Gland")
binbounds <- levels(cut(pcs.all$PC1[pcs.all$PC1<50],4))
binbounds <- unlist(lapply(strsplit(binbounds,","),function(x) x[2]))
binbounds <- as.numeric(gsub("]","",binbounds))
binbounds <- c(min(pcs.all$PC1),binbounds)

add <- data.frame("SampleID"=sce.sum$SampleID,
		  "Age"=paste0(sce.sum$Age,"w"),
		  "PC1"=-pcpredict[,1],
		  "PC2"=pcpredict[,2])

pcs.all$Age <- as.numeric(pcs.all$Age)

bins <- data.frame("PC1"=c(mean(c(binbounds[1],binbounds[2])),
			 mean(c(binbounds[2],binbounds[3])),
			 mean(c(binbounds[3],binbounds[4])),
			 mean(c(binbounds[4],binbounds[5])),
			 75),
		   "PC2"=c(60,60,60,60,60),
		   "Age"=c(mean(pcs.all$Age[pcs.all$PC1 < binbounds[2]]),
			   mean(pcs.all$Age[pcs.all$PC1 > binbounds[2] & pcs.all$PC1 < binbounds[3]]),
			   mean(pcs.all$Age[pcs.all$PC1 > binbounds[3] & pcs.all$PC1 < binbounds[4]]),
			   mean(pcs.all$Age[pcs.all$PC1 > binbounds[4] & pcs.all$PC1 < binbounds[5]]),
			   mean(pcs.all$Age[pcs.all$PC1 > binbounds[5]]))
		   )

bins$Age <- round(bins$Age,1)
bins$Age <- paste0(bins$Age,"w")


p1_col <- ggplot(pcs.all, aes(x=PC1, y=PC2, fill=Tumor)) +
    geom_rect(aes(xmin=-Inf,
		      xmax = binbounds[2],
		      ymin = -Inf,
		      ymax = Inf), fill = 'pink', alpha = 0.02) +
    geom_rect(aes(xmin=binbounds[2],
		      xmax = binbounds[3],
		      ymin = -Inf,
		      ymax = Inf), fill = 'pink', alpha = 0.06) +
    geom_rect(aes(xmin=binbounds[3],
		      xmax = binbounds[4],
		      ymin = -Inf,
		      ymax = Inf), fill = 'pink', alpha = 0.10) +
    geom_rect(aes(xmin=binbounds[4],
		      xmax = binbounds[5],
		      ymin = -Inf,
		      ymax = Inf), fill = 'pink', alpha = 0.20) +
    geom_rect(aes(xmin=binbounds[5],
		      xmax = Inf,
		      ymin = -Inf,
		      ymax = Inf), fill = 'red', alpha = 0.03) +
    geom_vline(xintercept=binbounds, lty="dashed") +
    geom_point(size=2,pch=21) + #This is just laziness 
    geom_point(data=add, size=4, pch=21, fill="blue") +
    geom_text(data=bins,aes(label=Age, fill=NULL),fontface="bold",size=6) +
    geom_text_repel(data=add,aes(label=Age, fill=NULL),fontface="bold",size=6) +
    theme_pub() +
    theme(legend.position="none") 
p1_col
```

# Session Info and Save
```{r, message=FALSE, warning=FALSE}
out <- pcs.all[,c("Condition","ptime")]
out.nt <- out[!grepl("TM",out$Condition),]
out.t <- out[grepl("TM",out$Condition),]
out.nt$ptimeBin <- as.numeric(cut(out.nt$ptime,4,labels=c(1,2,3,4)))
out.t$ptimeBin <- 5
out <- rbind(out.nt,out.t)
out$ptimeBin <- as.factor(out$ptimeBin)
write.csv(out,"../../data/Integrated/Robjects/TumorTime.csv")
sessionInfo()
```
