---
title: "Cell Calling for pregnancy dataset"
author: "Karsten Bach"
date: '`r Sys.Date()`'
output:
  html_notebook:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: journal
    highlight: tango
    code_folding: hide
---
***

# Call Empty Drops

I use the `emptyDrops` package to distinguish cells from backgrounds.
In the plots the below, the blue line indicates the inflection point and the black one the knee point.
Most profiles look quite nice.

```{r, warning=FALSE, message=FALSE}
library(DropletUtils)
library(scran)
library(BiocParallel)
library(Matrix)
library(ggplot2)
library(dplyr)
library(cowplot)
theme_set(theme_cowplot())

smps <- list.files("../../data/Pregnancy/CellRangerOutput/")
smps <- grep("[0-9]$",smps,value=TRUE)
fnames <- paste0("../../data/Pregnancy/PCellRangerOutput/",smps,"/raw")

out <- bplapply(fnames, function(fle) {
    nm <- gsub("../../data/Pregnancy/CellRangerOutput/|/raw","",fle)
    sce <- read10xCounts(fle, col.names=TRUE,version="3")

    bcrank <- barcodeRanks(counts(sce))

    # Only showing unique points for plotting speed.
    uniq <- !duplicated(bcrank$rank)
    fplot <- data.frame("rank"=bcrank$rank[uniq],
			"total"=bcrank$total[uniq]) 

    bcplt <- ggplot(fplot, aes(x=rank, y=total)) +
			geom_point() +
			scale_x_log10() +
			scale_y_log10() +
			geom_hline(yintercept=metadata(bcrank)$inflection,
				   lty="dashed",color="blue") +
			geom_hline(yintercept=metadata(bcrank)$knee,
				   lty="dashed") +
			ggtitle(nm) +
			xlab("Barcode Rank") +
			ylab("Total UMIs")


		    #     if (nm=="SampleC3") {

		    #     } else {
		    niter <- 10000
		    #     }

    set.seed(42)
    limit <- 100   
    e.out <- emptyDrops(counts(sce), lower=limit, niter=niter)

    tbl <- table(Sig=e.out$FDR <= 0.001, Limited=e.out$Limited)

    set.seed(42)
    all.out <- emptyDrops(counts(sce), lower=limit, test.ambient=TRUE, niter=niter)

    fplot <- data.frame("X"=all.out$PValue[all.out$Total <= limit & all.out$Total > 0])

    histplt <- ggplot(fplot, aes(x=X)) +
	geom_histogram(bins=20) +
	xlab("P-Value") +
	ggtitle(nm) 

    sce <- sce[,which(e.out$FDR<=0.001)]
    sce$Sample <- nm

    out <- list("SCE"=sce,
		"BCPLT"=bcplt,
		"TBL"=tbl,
		"HISTPLT"=histplt)

    print(paste0("Done with Sample ",nm))

    return(out)
}, BPPARAM=MulticoreParam(2))

scelist <- lapply(out, function(x) x[["SCE"]])
bcplts <- lapply(out, function(x) x[["BCPLT"]])
tbls <- lapply(out, function(x) x[["TBL"]])
histplts <- lapply(out, function(x) x[["HISTPLT"]])
gc()
rm(out)
```

## Some QC on cell calling

I here include some visualisation of the calling process as a QC step.

### Barcode Plots
```{r, warning=FALSE, message=FALSE, fig.width=13, fig.height=13}
plot_grid(plotlist=bcplts)
```

### P-Value Histograms
```{r, warning=FALSE, message=FALSE, fig.width=13, fig.height=13}
plot_grid(plotlist=histplts)
```

### P-Value limits
As the P-Values are not computed analytically you might have cells with P-Values that are 'limited' in their precision. If these are called non-significant, they might only be non-significant because of the lack of precision. 

- There are `r sum(unlist(lapply(tbls, function(x) x[1,2]))!=0)` samples that have cells with non-significant P-Values that are limited by the iterations. 


# Concatenate 
I then concatenated all samples into one SCE object.

```{r, warning=FALSE, message=FALSE}
addSampleID <- function(sce) {
    smp <- unique(sce$Sample)
    smp <- substring(smp,7,nchar(smp))
    bcs <- substring(colnames(sce),1,(nchar(colnames(sce))-1))
    colnames(sce) <- paste0(bcs,smp)
    return(sce)
}
scelist <- lapply(scelist,addSampleID)

m <- lapply(scelist,counts)
m <- do.call(cbind,m)
pD <- lapply(scelist,colData)
pD <- do.call(rbind,pD)
pD$barcode <- rownames(pD)
fD <- lapply(scelist,rowData)
stopifnot(all(unlist(lapply(fD[-1], function(x) identical(x,fD[[1]])))))
fD <- fD[[1]]
```

# Adding some other information to the SCE

I then added some meta data to the columns and row of the SCE.

## Column Data
```{r, warning=FALSE, message=FALSE}
# Read in meta data
meta.b1 <- read.csv("../../data/misc/SLX-18206.HKJ5GDMXX.s_2.contents.csv",stringsAsFactors=FALSE)
meta.b2 <- read.csv("../../data/misc/SLX-18386.HL3WKDMXX.s_1.contents.csv",stringsAsFactors=FALSE)
meta <- rbind(meta.b1,meta.b2)
meta$Barcode <- substr(meta$Barcode,5,6)
meta$Condition <- unlist(lapply(strsplit(meta$SampleName,"_"), function(x) x[1]))
meta$Batch <- ifelse(meta$SampleID %in% c("S13","S14","S15"),"2","1")

# Add more info to phenotype Data
pD <- data.frame(pD)
pD <- mutate(pD, SeqID=substr(barcode,18,nchar(barcode))) %>%
        mutate(SampleID=plyr::mapvalues(SeqID,meta$Barcode,
				  meta$SampleName)) %>%
	mutate(Condition=plyr::mapvalues(SampleID,meta$SampleName,
			     meta$Condition),
	       Batch=plyr::mapvalues(SampleID,meta$SampleName,
				     meta$Batch))
    
head(pD)
```

## Feature Data

```{r, warning=FALSE, message=FALSE}
# Add more info to the feature Data
mitoGenes <- read.table("../../data/misc/MitoGenes.txt")

# Ribosmal genes
library(biomaRt)
gos <- c("GO:0003735","G0:0005840","GO:0015935","GO:0015934") # constituents of ribosome
ensembl  <- useMart("ensembl",dataset="mmusculus_gene_ensembl",host="useast.ensembl.org") # at the time of writing the ensembl.org had issues
gene.data <- getBM(attributes=c('ensembl_gene_id', 'go_id'), mart=ensembl)
gene.data <- gene.data[gene.data$go_id %in% gos,]

fD$Ribosomal <- fD$ID %in% gene.data$ensembl_gene_id

#Add ribosomal proteins that have been missed
fD$Ribosomal <- fD$Ribosomal | grepl("Rpl",fD$Symbol)

# Mitochondrial
fD$Mitochondrial <- fD$ID %in% mitoGenes$V1

# Keep for hvg
fD$KeepForHvg <- !(fD$Ribosomal | fD$Mitochondrial)

fD$uniq <- scater::uniquifyFeatureNames(fD$ID,fD$Symbol)
rownames(fD) <- fD$uniq
stopifnot(identical(fD$ID,rownames(m)))
rownames(m) <- fD$uniq
head(fD)
```

## Save
Saving all data (for genes that have a total count higher than 1 accross the dataset)
```{r, warning=FALSE, message=FALSE}
sceout <- SingleCellExperiment(assays=list(counts=m),
			       colData=pD,
			       rowData=fD)
sceout <- sceout[,sceout$Condition!="11aOVX"]

keep <- rowSums(counts(sceout))>1
sceout <- sceout[keep,]
saveRDS(sceout,file="../../data/Pregnancy/Robjects/SCE.rds")
```

## Session Info
```{r}
sessionInfo()
```
