---
title: "CellType Annotation for Pregnancy Timecourse"
author: "Karsten Bach"
date: '`r Sys.Date()`'
output:
  html_notebook:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: united
    highlight: tango
    code_folding: hide
---
***

This annotation is still a bit rough and mainly aims at removing clusters in which I have low confidence so I can have a meaningful MNN mapping with the other dataset

```{r, message=FALSE,warning=FALSE}
library(scran)
library(scater)
library(plyr)
library(dplyr)
library(knitr)
library(ggplot2)
library(cowplot)
library(Matrix)
library(irlba)
source("../functions.R")
theme_set(theme_cowplot())

sce <- readRDS("../../data/Pregnancy/Robjects/SCE_QC_norm.rds")

umap <- read.csv("../../data/Pregnancy/Robjects/UMAP_corrected.csv",stringsAsFactors=FALSE)[,-1]
clst <- read.csv("../../data/Pregnancy/Robjects/Clusters_stepsize6.csv",stringsAsFactors=FALSE)[,-1]
anno <- left_join(umap,clst)
sce <- sce[,anno$barcode]
pD <- data.frame(colData(sce))
pD <- left_join(pD,anno[,c("barcode","UMAP1","UMAP2","Cluster")])
colData(sce) <- DataFrame(pD)
reducedDim(sce, type="UMAP") <- pD[,c("UMAP1","UMAP2")]
```


# Major Groups
```{r, fig.width=12, fig.height=4, warning=FALSE, message=FALSE}
# Epithelial
plotUMAP(sce, colour_by="Epcam", text_by="Cluster") + ggtitle("Epithelial")
plotExpression(sce, x="Cluster", features=c("Epcam")) + ggtitle("Epithelial")
epithelial <- c(2,26,28,30,37,40,47, # basal
		10,11,24,27,3,36,4,43,44,5)
epiCluster <- paste0("C",epithelial)

# Immune
plotExpression(sce, x="Cluster", features=c("Ptprc")) + ggtitle("Immune")
plotExpression(sce, x="Cluster", features=c("Cpa3")) + ggtitle("Immune")
plotExpression(sce, x="Cluster", features=c("Cd79a")) + ggtitle("Immune")
plotExpression(sce, x="Cluster", features=c("S100a8")) + ggtitle("Immune")
immune <- c(17,21,46,48,6,38,41,
	    12,13,20,33,35,7,
	    50, # Mast cells CPA3
	    14) # B-Cells Cd79a

neutros <- c(45) # not ptprc pos but S100a8
immuneCluster <- paste0("C",c(immune,neutros))

# Fibroblasts
plotExpression(sce, x="Cluster", features="Col3a1") + ggtitle("Fibroblasts")
fibroblasts <- c(15,16,25,29,31,34,39,49,8,9)
fibroblastsCluster <- paste0("C",fibroblasts)

mark <- findMarkers(sce,sce$Cluster,pval.type="all")


# Stroma
# 18 is Schwann Cell (Mbp, Mpz)
# 1 lymphatic endothelium (Nts, # not really Lyve1 but Nts marked the lymphatic endothelium in prev)
# 1,22,32 are endothelium (Emcn)
stroma <- c(1,22,32,# Endothelium 1 is lymphatic
	    18, # Schwann 
	    23, # Not sure what these are?!? Might not be stroma
	    19) # Pericytes
stromaCluster <- paste0("C",stroma)

plotExpression(sce, x="Cluster", features=c("Mpz")) + ggtitle("Schwann")
plotExpression(sce, x="Cluster", features=c("Emcn")) + ggtitle("Endothelium")
plotExpression(sce, x="Cluster", features=c("Nts")) + ggtitle("Endothelium")
plotExpression(sce, x="Cluster", features=c("Notch3")) + ggtitle("Pericytes")
plotExpression(sce, x="Cluster", features=c("Rgs5")) + ggtitle("Pericytes")

sce$MajorGroups <- mapvalues(as.character(sce$Cluster),
		       from=c(epiCluster, immuneCluster, fibroblastsCluster, stromaCluster),
		       to=c(rep("Epithelial",length(epiCluster)),
			    rep("Immune",length(immuneCluster)),
			    rep("Fibroblast",length(fibroblastsCluster)),
			    rep("Stroma",length(stromaCluster))))
# Finer Annotation
myeloid <- c(12,13,20,33,35,7,50,45)
lymphoid <- c(17,21,46,48,6,14,38,41)
basal <- c(2,26,28,30,37,40,47)
luminal <- c(10,11,24,27,3,36,4,43,44,5)

luminal <- paste0("C",luminal)
basal <- paste0("C",basal)
myeloid <- paste0("C",myeloid)
lymphoid <- paste0("C",lymphoid)
# unknown <- "C36"

sce$Groups <- mapvalues(as.character(sce$Cluster),
		       from=c(lymphoid, myeloid, basal, luminal, fibroblastsCluster, stromaCluster),
		       to=c(rep("Lymphoid",length(lymphoid)),
			    rep("Myeloid",length(myeloid)),
			    rep("Basal",length(basal)),
			    rep("Luminal",length(luminal)),
			    rep("Fibroblast",length(fibroblastsCluster)),
			    rep("Stroma",length(stromaCluster))))
			    
```

```{r, fig.width=7, fig.height=7, warning=FALSE, message=FALSE}
plotUMAP(sce, colour_by="Groups", text_by="Cluster")
```

# Cell Type mapping
```{r, fig.width=7, fig.height=7, warning=FALSE, message=FALSE}
celltypes <- c(1:50)
celltypes <- paste0("C",celltypes)

names(celltypes) <- c("Ec1","Bsl1","Lp","Hsp","HsGearly",#5
		      "Nk1","Dc209","Fb1","Fb2","Avp",#10
		      "Avd","Mp2a?","Mp1b?","Bcs","Uk1?",#15
		      "Fb3", "Nk2", "Swc", "Pcs1", "Mp1a",#20
		      #
		      "Tc.1Cyl","Ec2","Pc2","Hs","Fb4",#25
		      "BslNP","HsG","BslG2","Fb5","BslG3",#30
		      "Fb6", "FbUk?", "Dcs?", "Fb7", "Mp2b",#35
		      "Hs2", "BslCycl", "TcLN", "FbUk2?", "BslG4",#40
		      #
		      "TcLN2","Uk2?","HsCycl","AvpDbl?", "Np",#45
		      "Ilc3.1", "Dbl?", "Ilc3.2", "Fb8", "Mst" #50
		      )
sce$CellTypes <- mapvalues(sce$Cluster, from=celltypes, to=names(celltypes))
sort(table(sce$CellTypes), decreasing=TRUE)
```

# QC Check
- As this will be the final annotation, I will one last time check that none of the clusters is funny in terms of QC criteria

```{r, fig.width=10, fig.height=3, warning=FALSE, message=FALSE}
extraQC <- read.csv("../../data/Pregnancy/Robjects/QC_Part2.csv")[,-1]
rownames(extraQC) <- extraQC$barcode
extraQC <- extraQC[sce$barcode,]
sce$DbltScore <- extraQC$DbltScore
sce$IsPotDamaged <- extraQC$IsPotDamaged

plotColData(sce, x="CellTypes",y="UmiSums") + scale_y_log10() + theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))
plotColData(sce, x="CellTypes",y="GenesDetected") + scale_y_log10() + theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))
plotColData(sce, x="CellTypes",y="prcntMito") + theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))
plotColData(sce, x="CellTypes",y="DbltScore") + theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))
```

## Dbl? is a multiplet
- Expresses basal/luminal Krts and Ptprc
```{r, fig.width=12, fig.height=3, warning=FALSE, message=FALSE}
plotExpression(sce, "Krt8", x="CellTypes")
plotExpression(sce, "Krt17", x="CellTypes")
```

## Fb6 is not a multiplet
- receives high doublet scores
- marked by Apod
```{r, fig.width=12, fig.height=3, warning=FALSE, message=FALSE}
plotExpression(sce, "Apod", x="CellTypes")
```

## AvpDbl? is a multiplet
- Expresses basal/luminal Krts (see above)

## Clean Up
- Clean Up, removing the doublets and damaged cells
- I am also removing Uk1 and Uk2, neither of the clusters express specific markers, have v low count and don't map to anything else in the tumorigenesis set
- I am also removing TcLN and TcLN2 which only appear in CTRL_3 and I presume are contaminations from a lymphnode that wasn't removed properly.
- In addition they express either Ccr7 or Sell(Cd62l) markers for LN T-Cells.
```{r, fig.width=13, fig.height=3, warning=FALSE, message=FALSE}
tbbl <- table(sce$SampleID,sce$CellTypes)
tbbl.prop <- round(t(t(tbbl)/colSums(tbbl)) *100,2)
plotExpression(sce, x="CellTypes", features=c("Sell"))
plotExpression(sce, x="CellTypes", features=c("Ccr7"))
rmClusters <- c("AvpDbl?","Dbl?","Uk1?","Uk2?","TcLN","TcLN2")
sce <- sce[,!sce$CellTypes %in% rmClusters]
```


# Final Out
- Keep in mind that Tc.1 and Tc.2 are highly specific to the CTRL_3 sample, might need double checking
```{r, fig.width=7, fig.height=7, warning=FALSE, message=FALSE}
colnames(sce) <- sce$barcode
plotUMAP(sce, colour_by="Groups", text_by="CellTypes")
saveRDS(sce,file="../../data/Pregnancy/Robjects/SCE_final.rds")
```
