---
title: "Doublet Calling"
author: "Karsten Bach"
date: '`r Sys.Date()`'
output:
  html_notebook:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: journal
    highlight: tango
    code_folding: hide
---
***

In this document I will further clean up the data and remove cells that most likely represent doublets, damaged cells or red blood cell contamination.


# Removal of doublets
The following strategy for doubled removal was employed, somewhat following an approach from __Pijuana-Sala, Griffiths and Guibentif et al.__

1. Estimate probability of cell being a doublet per sample using doubletCells from scran.
2. Compute per sample clusters based on all genes based on a SNNGraph (k=10) and walktrap clustering with high resolution (steps=2, see `05_computeDoubletScores.R`).
3. Clusters with high average doublet probability are identified as doublets.
4. All cells and samples (incl. doublets) are integrated using fastMNN, cells that cluster with "known" doublets cells are guilty by association.

## Per sample clusters
- Below are the per sample clusters as well as the doublet scores per cell.
- The big clusters that all receive relatively high doublet scores are fibroblasts.
- Interestingly, CTRL_1 has this funny ring structure that Pauline had seen in the combined data set as well.
```{r, message=FALSE,fig.height=17,fig.width=17}
# Identifying doublets with doubletCluster
library(ggplot2)
library(scran)
library(cowplot)
library(wesanderson)
library(viridis)
theme_set(theme_cowplot())

# Read in data
ddata <- readRDS("../data/Robjects/DoubletData.rds")

# Add sample index to data frames
# Also I am log transforming the scores here already
# Since Bioc10 the scores have a different distribution, with a much wider range
nms <- names(ddata)
ddata <- lapply(names(ddata), function(x) {
		dfs <- ddata[[x]]
		dfs$Sample <- x
		dfs$DbltScore <- log10(dfs$DbltScore)
		return(dfs)
})
names(ddata) <- nms

# Visualizing the clustering
cluster_plots <- lapply(names(ddata), function(nm) {
		   x <- data.frame(ddata[[nm]])
		   p <- ggplot(x, aes(x=SubUMAP1, y=SubUMAP2, color=factor(Cluster))) +
			       geom_point(size=0.7) +
			       scale_color_manual(values=wes_palette("FantasticFox1",length(unique(x$Cluster)),type="continuous")) +
			       theme(legend.position="none", panel.background = element_rect(colour = "black", linetype = 1, size = 0.5),
				     axis.line = element_blank(),
				     axis.ticks = element_blank(),
				     axis.text = element_blank(),
				     axis.title = element_blank(),) +
			       ggtitle(paste0(nm," (n=",nrow(x),", k=",length(unique(x$Cluster)),")")) 
			   return(p)
})
names(cluster_plots) <- names(ddata)
plot_grid(plotlist=cluster_plots)

# Visualizing the doublet scores
score_plots <- lapply(names(ddata), function(nm) {
		   x <- data.frame(ddata[[nm]])
		   p <- ggplot(x, aes(x=SubUMAP1, y=SubUMAP2, color=DbltScore)) +
			       geom_point(size=0.7) +
			       scale_color_viridis() +
			       theme(legend.position="none", panel.background = element_rect(colour = "black", linetype = 1, size = 0.5),
				     axis.line = element_blank(),
				     axis.ticks = element_blank(),
				     axis.text = element_blank(),
				     axis.title = element_blank(),) +
			       ggtitle(paste0(nm," (n=",nrow(x),", k=",length(unique(x$Cluster)),")")) 
			   return(p)
})
names(score_plots) <- names(ddata)
plot_grid(plotlist=score_plots)
```

## Identify clusters per samples that represent doublets
- Clusters that had a higher than median + 1.5xmad doublet score and were present at a frequency of less than 5% were identified as doublet cells
- 1.5 as this seemed to produce the most reasonable results in terms of stringency
- At this point I want to be rather stringent in the definition as it is more problematic to remove biologically relevant clusters than maintaining doublet clusters as they can still be flagged later.
```{r, message=FALSE, fig.width=10}
# Identifying doublets based on doubletCell

ddatadf <- data.frame(do.call(rbind,ddata))

# Compute median scores per cluster and sample
medscores <- aggregate(ddatadf$DbltScore, list(ddatadf$Sample, ddatadf$Cluster), median)
names(medscores) = c("sample", "cluster", "median.score")

medscores$n.cells = sapply(1:nrow(medscores), function(row){
			       sum(ddatadf$Cluster == medscores$cluster[row] & ddatadf$Sample == medscores$sample[row])
})

medscores$fracCells = sapply(1:nrow(medscores), function(row){
				  medscores$n.cells[row]/sum(medscores$n.cells[medscores$sample == medscores$sample[row]])
})


# Define doublets based on fraction and median doublet score
medscores$outlier <- scater::isOutlier(medscores$median.score,nmad=1.5,type="higher",log=FALSE)
medscores$isDoublet <- medscores$outlier & medscores$fracCells <= 0.05

# Plot from Johnny with the significant clusters highlighted
ggplot(medscores, aes(x = sample, y = median.score, col = factor(cluster))) +
  geom_point(data = medscores[medscores$isDoublet,]) +
  geom_jitter(data = medscores[!medscores$isDoublet,], col = "darkgrey", size = 0.4, width = 0.2, height = 0) +
  guides(color="none") +
  labs(x = "Sample", y = "cluster median score") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

medscores.dbls <- medscores[medscores$isDoublet,]
smp.dbltCluster <- paste(medscores.dbls$sample,medscores.dbls$cluster,sep="_")
```


## Visualizing the doublet called clusters
- most clusters show the typical tear shape in the UMAP
- some clusters also appear to bel self-doublets, which are less problematic but will still be removed
- however we can already see here that there might be more doublet clusters than we identified above
```{r, message=FALSE, fig.height=13, fig.width=13}
smps <- names(ddata)[names(ddata) %in% medscores.dbls$sample]
doublet_plots <- lapply(smps, function(nm) {
		   x <- data.frame(ddata[[nm]])
		   p <- ggplot(x, aes(x=SubUMAP1, y=SubUMAP2)) +
			       geom_point(color="grey80",size=0.7) +
			       geom_point(data=x[x$Cluster %in% medscores.dbls[medscores.dbls$sample==nm,"cluster"],], aes(color=factor(Cluster))) +
			       #                                scale_color_manual(values=wes_palette("FantasticFox1",length(unique(x$Cluster)),type="continuous")) +
			       theme(legend.position="none", panel.background = element_rect(colour = "black", linetype = 1, size = 0.5),
				     axis.line = element_blank(),
				     axis.ticks = element_blank(),
				     axis.text = element_blank(),
				     axis.title = element_blank(),) +
			       ggtitle(paste0(nm," (n=",nrow(x),", k=",length(unique(x$Cluster)),")")) 
			   return(p)
})
names(doublet_plots) <- smps
plot_grid(plotlist=doublet_plots)
```

## Doublet clusters have a larger library size
```{r, message=FALSE}
pD <- colData(readRDS("../data/Robjects/SCE_QC_norm.rds"))
pD <- data.frame(pD)
library(dplyr)
ddatadf <- left_join(data.frame(ddatadf),pD[,c("barcode","UmiSums","GenesDetected","prcntMito")])
ddatadf$sampClust <- paste0(ddatadf$Sample,"_",ddatadf$Cluster)
ddatadf$isDoublet <- ifelse(ddatadf$sampClust %in% smp.dbltCluster,"Doublet","Singlet")
ggplot(ddatadf, aes(x=isDoublet,y=log10(UmiSums))) +
       geom_boxplot() 
```

## Cluster batch corrected data
- Next we will identify cells that are guilty-by-association by clustering the full dataset including the above identified doublet cells and find clusters with a high frequency of doublets
```{r, message=FALSE, warning=FALSE}
# Read in BC PCA
sce.cor <- readRDS("../data/Robjects/Corrected_preDoublet.rds")
m.cor <- reducedDim(sce.cor, type="corrected")
igr <- buildSNNGraph(m.cor,d=NA, transposed=TRUE)
# Overcluster
cl <- igraph::cluster_walktrap(igr,steps=2)
cluster <- cl$membership
clusDf <- data.frame("barcode"=rownames(m.cor),
		     "fullCluster"=cluster)
ddatadf <- dplyr::left_join(ddatadf,clusDf)
```


## Guilty-by-association clusters
- All clusters with more than median+ 2x mad fraction of doublets are additionally identifed as doublet clusters
```{r, message=FALSE, warning=FALSE}
smry <- group_by(ddatadf, fullCluster, isDoublet) %>%
    summarize(NType=n()) %>%
    mutate(Prop=NType/sum(NType))

doubDef <- scater::isOutlier(smry$Prop[smry$isDoublet=="Doublet"],nmads=2,type="higher")
dblClusters <- smry$fullCluster[smry$isDoublet=="Doublet"][doubDef]
thrsld <- attributes(doubDef)[[1]][2]

ggplot(smry, aes(x=fullCluster, y=Prop, fill=isDoublet)) +
    geom_bar(stat="identity") +
    geom_hline(yintercept=1-thrsld) +
    scale_fill_manual(values=c("black","red"))

ump <- reducedDim(sce.cor, type="umap")
ump <- data.frame("barcode"=rownames(ump),
		  "UMAP1"=ump[,1],
		  "UMAP2"=ump[,2])
ddatadf <- left_join(data.frame(ddatadf),ump)

ggplot(ddatadf, aes(x=UMAP1, y=UMAP2, color=ddatadf$fullCluster %in% dblClusters)) +
    geom_point() +
    scale_color_manual(values=c("black","red")) +
    theme(legend.position="none")
```

## Highlighting all cells identified as doublets
```{r, message=FALSE}
dbl.barcodes <- ddatadf$barcode[ddatadf$isDoublet=="Doublet" | ddatadf$fullCluster %in% dblClusters]

ddatadf$isDoubletFinal <- ddatadf$barcode %in% dbl.barcodes
ggplot(ddatadf, aes(x=UMAP1, y=UMAP2, color=ddatadf$isDoubletFinal)) +
    geom_point() +
    scale_color_manual(values=c("black","red")) +
    theme(legend.position="none")
```

- As expected sample with more cells tend to have a higher frequency of doublets.
```{r, message=FALSE}
smry <- group_by(ddatadf, Sample, isDoubletFinal) %>%
    summarize(n=n()) %>%
    mutate(DoubletFraction=n/sum(n),totCells=sum(n)) %>%
    ungroup() %>%
    filter(isDoubletFinal)

ggplot(smry, aes(x=totCells,y=DoubletFraction)) +
    geom_point()
```

# Other outliers
- To try to identify damaged cells I looked at % of mitochondrial reads versus library size
- The idea being that clusters with very low mitochondrial reads can be damaged cells/ stripped nuclei
- Although there are some clsuters with low % of mito reads it is hard to visually define a threshold here and clusters below the threshold will only be flagged

```{r, message=FALSE, warning=FALSE}
m <- logcounts(readRDS("../data/Robjects/SCE_QC_norm.rds"))
hbb.exp <- colSums(m[c("Hbb-bt","Hbb-bs","Hbb-bh1"),])
hbb.exp <- data.frame("barcode"=names(hbb.exp),
		      "hbbExp"=hbb.exp)
ddatadf <- left_join(ddatadf,hbb.exp)
smry <- group_by(ddatadf, sampClust) %>%
    summarize(medUmi=median(UmiSums),
	      medprcntMito=median(prcntMito),
	      medGd=median(GenesDetected),
	      medHb=median(hbbExp)) %>%
    ungroup() %>%
    as.data.frame()

ggplot(smry, aes(x=medUmi,y=medprcntMito)) +
    geom_point() +
    scale_x_log10() +
    scale_y_log10() +
    geom_hline(yintercept=0.003) +
    xlab("Library Size") +
    ylab("Fraction of Mitochondrial Reads")
```

- Also in the UMAP below I am not fully convinced that they are all truly damaged cells.
- If they were I would expect them to "stand out" more. (plus it's only a handful of cells)
- I did label cells as "potentially damaged" and will keep that information for further analyses. 

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
cls <- smry[smry$medprcntMito<0.003,"sampClust"]
smps.dmgd <- unlist(lapply(strsplit(cls,split="_"),function(x) paste0(x[1],"_",x[2])))
cls.dmgd <- unlist(lapply(strsplit(cls,split="_"),function(x) x[3]))

dmgd_plots <- lapply(1:length(smps.dmgd), function(i) {
		   nm <- smps.dmgd[i]
		   x <- data.frame(ddata[[nm]])
		   p <- ggplot(x, aes(x=SubUMAP1, y=SubUMAP2)) +
			       geom_point(color="grey80",size=0.7) +
			       geom_point(data=x[x$Cluster %in% cls.dmgd[i],], aes(color=factor(Cluster))) +
			       #                                scale_color_manual(values=wes_palette("FantasticFox1",length(unique(x$Cluster)),type="continuous")) +
			       theme(legend.position="none", panel.background = element_rect(colour = "black", linetype = 1, size = 0.5),
				     axis.line = element_blank(),
				     axis.ticks = element_blank(),
				     axis.text = element_blank(),
				     axis.title = element_blank(),) +
			       ggtitle(paste0(nm," (n=",nrow(x),", k=",length(unique(x$Cluster)),")")) 
			   return(p)
})
names(dmgd_plots) <- smps.dmgd
plot_grid(plotlist=dmgd_plots)
dmgd <- ddatadf$barcode[ddatadf$sampClust %in% cls]
ddatadf$isPotDamaged <- ddatadf$barcode %in% dmgd
```

- Finally I exclude clusters that show expression of hemoglobin
```{r, message=FALSE, warning=FALSE}
library(ggrepel)
ggplot(smry, aes(x=sampClust,y=medHb)) +
    geom_point() +
    xlab("Per-Sample Clusters") +
    ylab("Median Hbb Expression") +
    geom_label_repel(data=smry[smry$medHb>0.01,],aes(label=sampClust)) +
    theme(axis.text.x=element_blank())
rbcs <- smry[smry$medHb>0.01,"sampClust"]
rbcs <- ddatadf$barcode[ddatadf$sampClust %in% rbcs]
```
# Save
```{r, message=FALSE, warning=FALSE}
ddatadf$isRbc <- ddatadf$barcode %in% rbcs
qcMets <- ddatadf[,c("barcode","DbltScore","isRbc","isDoubletFinal","isPotDamaged")]
write.csv(qcMets,"../data/Robjects/QC_Part2.csv")
```
