---
title: "Lineage Tracing Quality Control Report"
author: "Karsten Bach"
date: '`r Sys.Date()`'
output:
  html_notebook:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: united
    highlight: tango
---
***

# Load Data
```{r, message=FALSE}
library(scran)
library(scater)
library(dplyr)
library(knitr)
library(ggplot2)
library(Rtsne)
library(cowplot)
library(Matrix)
# source("functions.R")

# Load Data
dataList <- readRDS("../data/Robjects/ExpressionList.rds")
m <- dataList[["counts"]]
pD <- dataList[["phenoData"]]
fD <- dataList[["featureData"]]
rm(dataList)
```

# Number of Cells
We sequenced slightly less cells than expected, with a total of `r ncol(m)` cells.
```{r}
kable(t(as.matrix(table(pD$SampleID))))
kable(t(as.matrix(table(pD$Condition))))
```

## High reproducibility between technical replicates
```{r}
fp <- group_by(pD[!pD$AnimalHadTumor,], SampleID) %>%
    summarize(Number=n())
fp$Condition <- as.factor(substring(fp$SampleID,1,nchar(fp$SampleID)-1))
ggplot(fp, aes(x=Condition, y=Number)) +
    geom_point()
```

# Quality Control
```{r}
# Sequencing Depth and Genes detected
pD$UmiSums<- colSums(m)
pD$GenesDetected <- colSums(m!=0)
genesDetected <- ggplot(pD, aes(x=SampleID,y=GenesDetected,fill=Condition)) +
    geom_violin(draw_quantiles=0.5)+
    scale_y_log10() +
    ylab("Total number of genes detected") +
    ggtitle("Number of Genes detected") +
    theme_bw()
LibrarySize <- ggplot(pD, aes(x=SampleID,y=UmiSums,fill=Condition)) +
    geom_violin(draw_quantiles=0.5)+
    scale_y_log10() +
    ylab("Total number of molecules") +
    ggtitle("Number of Molecules") +
    theme_bw()

# Genewise QC
ntop <- 50 
mRel <- t(t(m)/colSums(m))
rownames(mRel)  <- fD$symbol
topExpressed <- apply(mRel,1,median)
names(topExpressed) <- rownames(mRel)
topExpressed <- topExpressed %>% sort(.,decreasing=TRUE) %>% names
plotData <- as.matrix(t(mRel)[,topExpressed[1:ntop]]) %>%
    reshape2::melt() %>%
    rename(Cell=Var1,
	   Gene=Var2,
	   RelativeExpression=value)
topGenes <- ggplot(plotData, aes(x=Gene, y=RelativeExpression)) +
    geom_boxplot() +
    coord_flip() +
    ggtitle("Most highly expressed genes") +
    theme_bw()
freqOfExp <- m!=0
rownames(freqOfExp) <- fD$symbol
freqOfExp <- sort(rowSums(freqOfExp)/ncol(freqOfExp),decreasing=TRUE)
plotData <- data.frame("Gene"=names(freqOfExp),"Frequency"=freqOfExp)
topFreq <- ggplot(plotData[1:ntop,], aes(x=factor(Gene,levels=Gene),y=Frequency)) +
    geom_bar(stat="identity") +
    coord_flip() +
    xlab("Gene") +
    ggtitle("Most frequently expressed genes") +
    theme_bw()

# Cell Viability
mMito <- m[fD$Mitochondrial,]
idtop <- fD[fD$symbol %in% names(freqOfExp)[1:ntop],"id"]
mTop <- m[idtop,]!=0
pD$prcntTop <- colSums(mTop)/ntop
pD$prcntMito <- colSums(mMito)/colSums(m)
cellViability <- ggplot(pD, aes(x=prcntMito, y=GenesDetected, color=Condition, shape=Replicate))+
    geom_point() +
    xlab("Percentage of mitochondrial reads") +
    ggtitle("Cell Viability") 
cellViability
rm(mMito)
rm(mTop)

genesDetected
LibrarySize
cellViability
topGenes
topFreq
sumry <- group_by(pD, SampleID) %>%
    summarize("Number of cells"=n(),
	      "Total molecules"=median(UmiSums),
	      "Genes Detected"=median(GenesDetected))
kable(sumry,caption="Summary of QC parameters as median")
```

# Index Swapping
```{r, eval=FALSE}
barcodes <- as.character(pD$barcode)
spt <- strsplit(barcodes, split = "-", fixed = T)
pD$sample <- sapply(spt, function(x) x[2])
pD$bcs <- sapply(spt, function(x) x[1])
pD.add <- data.frame(bcs=names(table(pD$bcs)),
		     bc.obs=(table(pD$bcs)))
pD.add <- pD.add[,-2]
pD <- dplyr::left_join(pD,pD.add)

# P1
index.p1 <- ggplot(pD, aes(x=SampleID, fill=as.factor(bc.obs.Freq))) +
    geom_bar() +
    #     ggtitle("Samples contain many shared barcodes") +
    scale_fill_discrete(name="Times barcode observed") +
    theme(legend.position="bottom",
	  legend.direction="horizontal")

compare <- function(barcodes, samples) {
    out <- NULL
    ids <- levels(samples)
    combs <- combn(ids,m=2, simplify=FALSE)
    for (i in seq_along(combs)) {
	comb <- combs[[i]]
	s1 <- comb[1]
	s2 <- comb[2]
	bc1 <- as.character(barcodes[samples==s1])
	bc2 <- as.character(barcodes[samples==s2])
	x <- length(intersect(bc1,bc2))
	m <- length(bc1)
	n <- 750000
	k <- length(bc2)
	p.val <- phyper(q=x-1,m=m,n=n,k=k,lower.tail=FALSE)
	tmp <- data.frame(s1=s1,
			  s2=s2,
			  n1=m,
			  n2=k,
			  shared=x,
			  p.val=p.val)
	out <- rbind(out,tmp)
    }
    return(out)
}

# P2
compDf <- compare(pD$bcs, as.factor(pD$SampleID))
index.p2 <- ggplot(compDf, aes(x=shared, y=-log10(p.val))) +
    geom_point() +
    xlab("# shared Barcodes") +
    ylab("-log10(P)") +
    geom_hline(yintercept=2,lty="dashed",color="red") 

index.p1
index.p2
```

# QC Filtering 
## Cell-wise
The following filters were applied to remove low quality cells:
    - Library size (3 mads from median)
    - Number of Genes detected (3 mads from median)
    - Mitochondrial reads (3 mads from median)
    - In addition cells with shared barcodes were also removed (to prevent reads from index swapping to contaminate the dataset)

```{r}
lib.outlier <- isOutlier(pD$UmiSums,nmads=3,type="both",log=TRUE)
gd.outlier <- isOutlier(pD$GenesDetected,nmads=3,type="both",log=TRUE)
mt.outlier <- isOutlier(pD$prcntMito,nmads=3,type="higher")
bc.shared <- pD$bc.obs.Freq>1

pD$PassLibSize <- !lib.outlier
pD$PassGenesDetected <- !gd.outlier
pD$PassViability <- !mt.outlier
# pD$PassIndexSwap <- !bc.shared
pD$PassAll <- pD$PassLibSize & pD$PassGenesDetected & pD$PassViability 
kable(table(pD$PassLibSize,pD$Condition))
kable(table(pD$PassGenesDetected,pD$Condition))
kable(table(pD$PassViability,pD$Condition))
kable(table(pD$PassAll,pD$Condition))
```

Overall a relatively high number of cells is flagged as "poor-quality", this is mainly due to a stringent filter on cell viability with `r sum(mt.outlier)` cells being flagged as non-viable.
The thresholds are as follows : `r max(pD$UmiSums[!lib.outlier])` for UmiSums,`r max(pD$GenesDetected[!gd.outlier])` for genes detected and `r max(pD$prcntMito[!mt.outlier])` for viability.
Yet, there does not seem to be a certain sample/Condition that has a substantially high number of low quality cells.

```{r}
kable(t(table(pD$PassAll,pD$SampleID))/colSums(table(pD$PassAll,pD$SampleID)))
kable(t(table(pD$PassAll,pD$Condition))/colSums(table(pD$PassAll,pD$Condition)))
```

## Gene-wise
Genes with a mean log expression lower that 0.01 are removed from the dataset, apart form Cre and tdTomato
```{r}
fD$keep <- rowMeans(m)>0.01 
```

## Final data & Save
```{r}
dim(m[fD$keep,pD$PassAll])
pD.pass <- pD[pD$PassAll,]
genesDetected %+% pD.pass
LibrarySize %+% pD.pass
cellViability %+% pD.pass
out <- list()
out[["counts"]] <- m
out[["phenoData"]] <- pD
out[["featureData"]] <- fD
saveRDS(out,file="../data/Robjects/ExpressionList_QC.rds")
```
