---
title: "Testing for Diff Correlation with Csn2"
author: "Karsten Bach"
date: '`r Sys.Date()`'
output:
  html_notebook:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: united
    highlight: tango
    code_folding: hide
---
***

```{r, message=FALSE,warning=FALSE}
library(slingshot)
library(scran)
library(scater)
library(ggplot2)
library(cowplot)
library(Matrix)
library(umap)
library(dplyr)
library(viridis)
library(ggrastr)
source("../functions.R")
source("../TwoSampleDCARS.R")
theme_set(theme_cowplot())

sce.full <- readRDS("../../data/Integrated//Robjects/SCE_combined_final.rds")
sce.full <- sce.full[,sce.full$MajorGroups=="Epithelial"]

sce <- sce.full[,sce.full$Experiment=="Tumorigenesis"]
rD <- rowData(readRDS("../../data/Tumorigenesis/Robjects/SCE.rds"))
rmgenes <- rownames(rD)[!rD$KeepForHvg]
rmgenes <- rownames(sce)[rownames(sce) %in% rmgenes]
outdr <- "../../data/figures/Figure2/"
```

# DCARS
What genes are differentially correlated between Csn2 and everything else during gestation and tumorigenesis 
in Lps and Avd.

```{r, message=FALSE,warning=FALSE}
sce.lp <- sce.full[,sce.full$CellTypesFinal %in% c("Lp","Avd") & sce.full$Condition!="WTOld"]
cd <- data.frame(colData(sce.lp))
cd <- droplevels(cd)
colData(sce.lp) <- DataFrame(cd)
dcars <- readRDS("../../data/Downstream/DCARS_OUT.rds")
#dcars <- dcars[!dcars$gene2 %in% rmgenes,]
head(dcars[order(dcars$normFDR),],20)
dcars$Gene <- as.character(dcars$gene2)
dcars <- dcars[!dcars$Gene %in% rmgenes,]
library(schex)
sce.lp <- make_hexbin(sce.lp, nbins = 50, 
		   dimension_reduction = "UMAP", use_dims=c(1,2))

pltGEX <- function(sce.full, gene) {
    plot_hexbin_gene(sce.full, gene=gene, type="logcounts",
		     action="mean") +
	theme_void() +
	theme(legend.position="none") +
	scale_fill_viridis(option="C") +
	ggtitle(gene)
}
sce.p <- sce.lp[,sce.lp$Experiment=="Pregnancy"]
sce.t <- sce.lp[,sce.lp$Experiment=="Tumorigenesis"]
sce.p <- make_hexbin(sce.p, nbins = 50, 
		   dimension_reduction = "UMAP", use_dims=c(1,2))
sce.t <- make_hexbin(sce.t, nbins = 50, 
		   dimension_reduction = "UMAP", use_dims=c(1,2))

plot_grid(pltGEX(sce.p,"Cebpb"),pltGEX(sce.t, "Cebpb"))


fplot <- dcars
fplot$Gene <- unlist(lapply(strsplit(rownames(fplot),"_"), function(x) x[[2]]))
fplot <- fplot[order(fplot$normFDR),]

library(ggrepel)
ggplot(fplot[1:40,], aes(x=cors_first_obs, y=cors_second_obs)) +
    geom_point(data=fplot[,c("cors_first_obs", "cors_second_obs")], size=.1, color="grey30") +
    geom_point() +
    geom_label_repel(aes(label=Gene)) +
    geom_abline(slope=1) +
    xlab("Csn2 Correlation in Pregnancy") +
    ylab("Csn2 Correlation in Tumorigenesis") +
    coord_fixed()

tf <- read.table("~/Cambridge/MammaryGlandDev/repo/data/miscData/TFcheckpoint_WithENSID.tsv",
		 header=TRUE, sep="\t", stringsAsFactors=FALSE)
tfs <- rownames(rD[rD$ID %in% tf$ensembl_gene_id,])


topTfs <- fplot[,]
tfs <- c("Nfkb1","Rel","Cebpb","Bcl11a","Sox10","Foxc1","Elf5","Foxq1","Foxl1",
	 "Barx2","Gata3","Gata6","Klf5","Ehf","Grhl1","Hey2")
topTfs <- topTfs[topTfs$Gene %in% tfs,]
library(ggrepel)
ggplot(topTfs, aes(x=cors_first_obs, y=cors_second_obs)) +
    geom_point(data=fplot[,c("cors_first_obs", "cors_second_obs")], size=.1, color="grey30") +
    geom_point() +
    geom_label_repel(aes(label=Gene)) +
    geom_abline(slope=1) +
    xlab("Csn2 Correlation in Pregnancy") +
    ylab("Csn2 Correlation in Tumorigenesis") +
    coord_fixed()
```

## Example Genes
```{r, message=FALSE,warning=FALSE}
cd$Csn2Expr <- logcounts(sce.lp)["Csn2",]
cd$Expr <- logcounts(sce.lp)["Lpl",]
ggplot(cd, aes(x=Csn2Expr,y=Expr)) +
    geom_point(size=.3,color="grey40") +
    geom_density_2d(adjust=6,color="black",lty="dashed") +
    geom_smooth(method="lm",color="black") +
    facet_wrap(~Experiment)

genes <- c("Cebpb","Olah","Cd36","Lpl","Acbd7","Lcn2","Eno1","Pdia6","Ndrg1","Atp6v1c2")
library(RColorBrewer)
col.t <- brewer.pal(n=6,"PuRd")[5]
col.p <- brewer.pal(n=5,"GnBu")[5]

for (gene in genes) {
    cd$Expr <- logcounts(sce.lp)[gene,]
    cor1 <- dcars[paste0("Csn2_",gene),"cors_first_obs"]
    cor1 <- sprintf("rho == %0.2f", cor1)
    cor2 <- dcars[paste0("Csn2_",gene),"cors_second_obs"]
    cor2 <- sprintf("rho == %0.2f", cor2)
    dmy <- data.frame(Csn2Expr=c(1.5,1.5),
		      Expr=c(max(cd$Expr)+.01,max(cd$Expr)+.01),
		      Experiment=c("Pregnancy","Tumorigenesis"),
		      Correlation=c(cor1,cor2))

    # Bandwidth for the density2d (set to non-zero if estimated zero)
    require(MASS)
    bws <- c(ifelse(bandwidth.nrd(cd$Csn2Expr) == 0, 0.1, bandwidth.nrd(cd$Csn2Expr)),
		ifelse(bandwidth.nrd(cd$Expr) == 0, 0.1, bandwidth.nrd(cd$Expr)))


    p0 <- ggplot(cd, aes(x=Csn2Expr,y=Expr)) +
        geom_rect(data = cd[cd$Experiment=="Pregnancy",][1,], aes(fill=Experiment), xmin=-Inf,xmax=Inf,
            ymin=-Inf,ymax=Inf, alpha=0.35) +
        geom_rect(data = cd[cd$Experiment=="Tumorigenesis",][1,], aes( fill = Experiment) ,xmin=-Inf,xmax=Inf,
            ymin=-Inf,ymax=Inf, alpha=0.35) +
	geom_point(size=.9, color="black") +
	geom_density_2d(color="grey95",lty="dashed",h=bws*6,lwd=1.2) +
	facet_wrap(~Experiment) +
        geom_smooth(method="lm",color="grey95",lwd=.8) +
	geom_text(data=dmy,aes(label=Correlation),size=7,fontface="bold",parse=TRUE) +
	theme(panel.grid.major=element_line(colour="grey80",size=0.1,linetype="dashed"),
	      axis.text=element_text(size=14),
	      axis.title=element_text(size=16,face="bold"),
	      #               strip.text.x = element_text(size=12, face="bold") ,
	      #               strip.text.y = element_text(size=12, face="bold"),
	      strip.text.x = element_blank() ,
	      strip.text.y = element_text(),
	      strip.background = element_rect(fill="white"),# colour="black"),
	      axis.line=element_line(size=1.5),
	      axis.ticks=element_line(size=1.5),
	      legend.position="none") +
	xlab("") +
	ylab("") +
	scale_fill_manual(values=c(col.p,col.t)) 
    p0

    ggsave(paste0("../../data/figures/Figure3/Csn2_", gene, ".png"),p0, width=7,height=3.95,dpi=600)
}
```



## Correcting the trend 
### Version 1 fit a trend and take residual
- Don't like this, this generates funky stuff on the extremes (e.g. Csn1s1, which has similar correlation in both)
```{r, message=FALSE,warning=FALSE}
ft <- lm(cors_second_obs ~ cors_first_obs, data=dcars)

ggplot(dcars, aes(x=cors_first_obs ,y=cors_second_obs)) +
    geom_point(size=.3) +
    geom_smooth(method="lm") 
dcars$Residual <- ft$residuals
fplot <- dcars[order(abs(dcars$Residual),decreasing=TRUE),]

library(ggrepel)
ggplot(fplot[1:40,], aes(x=cors_first_obs, y=cors_second_obs)) +
    geom_point(data=fplot[,c("cors_first_obs", "cors_second_obs")], size=.1, color="grey30") +
    geom_point() +
    geom_smooth(data=fplot[,c("cors_first_obs", "cors_second_obs")], method="lm") +
    geom_label_repel(aes(label=Gene)) +
    geom_hline(yintercept=0) +
    geom_vline(xintercept=0) +
    xlab("Csn2 Correlation in Pregnancy") +
    ylab("Csn2 Correlation in Tumorigenesis") +
    coord_fixed(1)
```
### Version 2 rescale to median 0

```{r, message=FALSE,warning=FALSE}
dcars$first_dist_to_med <- dcars$cors_first_obs - median(dcars$cors_first_obs)
dcars$second_dist_to_med <- dcars$cors_second_obs - median(dcars$cors_second_obs)
dcars$dist_med <-dcars$first_dist_to_med - dcars$second_dist_to_med

fplot <- dcars[order(abs(dcars$dist_med), decreasing=TRUE),]
fplot.sig <- fplot[fplot$normFDR<0.001 & abs(fplot$dist_med)>0.3,]
genes <- c(fplot.sig[order(fplot.sig$dist_med),][1:10,"Gene"],fplot.sig[order(-fplot.sig$dist_med),][1:10,"Gene"])
genes <- c(genes,"Cebpb","Olah","Glycam1")
ggplot(fplot.sig, aes(x=first_dist_to_med, y=second_dist_to_med)) +
    geom_point(data=fplot[,c("first_dist_to_med", "second_dist_to_med")], size=.1, color="grey30") +
    geom_point() +
    geom_hline(yintercept=0, lty="dashed", size=1.2) +
    geom_vline(xintercept=0, lty="dashed", size=1.2) +
    geom_label_repel(data=fplot.sig[fplot.sig$Gene %in% genes,], aes(label=Gene),size=6) +
    geom_abline(slope=1, lty="dashed",size=1.2, color="grey50") +
    xlab("Distance from median Csn2 correlation in Pregnancy") +
    ylab("Distance from median Csn2 Correlation in Tumorigenesis") +
    scale_x_continuous(breaks=c(-0.5,0,0.5)) +
    scale_y_continuous(breaks=c(-0.5,0,0.5)) +
    coord_fixed(1) +
    theme(axis.line=element_line(size=1.5),
	  axis.ticks=element_line(size=1.5),
	  axis.text=element_text(size=14),
	  axis.title=element_text(size=16))
ggsave("../../data/figures/Figure3/DCARS.pdf")



#Second interesting opposed cor in pregnancy
#Int1 is included in int2
fplot$int2 <- fplot$normFDR<0.01 & abs(fplot$dist_med)>0.3 & sign(fplot$first_dist_to_med)!=sign(fplot$second_dist_to_med)

ggplot(fplot[fplot$int1,], aes(x=first_dist_to_med, y=second_dist_to_med)) +
    geom_point(data=fplot[,c("first_dist_to_med", "second_dist_to_med")], size=.1, color="grey30") +
    geom_point() +
    geom_label_repel(aes(label=Gene)) +
    geom_abline(slope=1, lty="dashed") +
    geom_hline(yintercept=0) +
    geom_vline(xintercept=0) +
    xlab("Distance from median Csn2 correlation in Pregnancy") +
    ylab("Distance from median Csn2 Correlation in Tumorigenesis") +
    coord_fixed(1)
```

```{r, message=FALSE,warning=FALSE}
genes <- fplot.sig$Gene
library(RColorBrewer)
col.t <- brewer.pal(n=6,"PuRd")[5]
col.p <- brewer.pal(n=5,"GnBu")[5]

pltlist <- list()
for (gene in genes) {
    cd$Expr <- logcounts(sce.lp)[gene,]
    cor1 <- dcars[paste0("Csn2_",gene),"cors_first_obs"]
    cor1 <- sprintf("rho == %0.2f", cor1)
    cor2 <- dcars[paste0("Csn2_",gene),"cors_second_obs"]
    cor2 <- sprintf("rho == %0.2f", cor2)
    dmy <- data.frame(Csn2Expr=c(1.5,1.5),
		      Expr=c(max(cd$Expr)+.01,max(cd$Expr)+.01),
		      Experiment=c("Pregnancy","Tumorigenesis"),
		      Correlation=c(cor1,cor2))

    # Bandwidth for the density2d (set to non-zero if estimated zero)
    require(MASS)
    bws <- c(ifelse(bandwidth.nrd(cd$Csn2Expr) == 0, 0.1, bandwidth.nrd(cd$Csn2Expr)),
		ifelse(bandwidth.nrd(cd$Expr) == 0, 0.1, bandwidth.nrd(cd$Expr)))


    pltlist[[gene]] <- ggplot(cd, aes(x=Csn2Expr,y=Expr)) +
        geom_rect(data = cd[cd$Experiment=="Pregnancy",][1,], aes(fill=Experiment), xmin=-Inf,xmax=Inf,
            ymin=-Inf,ymax=Inf, alpha=0.35) +
        geom_rect(data = cd[cd$Experiment=="Tumorigenesis",][1,], aes( fill = Experiment) ,xmin=-Inf,xmax=Inf,
            ymin=-Inf,ymax=Inf, alpha=0.35) +
	geom_point(size=.9, color="black") +
	geom_density_2d(color="grey95",lty="dashed",h=bws*6,lwd=1.2) +
	facet_wrap(~Experiment) +
        geom_smooth(method="lm",color="grey95",lwd=.8) +
	geom_text(data=dmy,aes(label=Correlation),size=5,fontface="bold",parse=TRUE) +
	theme(panel.grid.major=element_line(colour="grey80",size=0.1,linetype="dashed"),
	      axis.text=element_text(size=14),
	      axis.title=element_text(size=16,face="bold"),
	      #               strip.text.x = element_text(size=12, face="bold") ,
	      #               strip.text.y = element_text(size=12, face="bold"),
	      strip.text.x = element_blank() ,
	      strip.text.y = element_text(),
	      strip.background = element_rect(fill="white"),# colour="black"),
	      axis.line=element_line(size=1.5),
	      axis.ticks=element_line(size=1.5),
	      legend.position="none") +
	xlab("") +
	ylab("") +
	ggtitle(gene) +
	scale_fill_manual(values=c(col.p,col.t)) 
}
plot_grid(plotlist=pltlist[120:137])
```

## Enrichment of top left corner
```{r, message=FALSE,warning=FALSE, eval=FALSE}


# fplot <- dcars
# library(ggrepel)
# ggplot(fplot[fplot$intTum,][1:30,], aes(x=cors_first_obs, y=cors_second_obs)) +
#     geom_point(data=fplot[,c("cors_first_obs", "cors_second_obs")], size=.1, color="grey30") +
#     geom_point() +
#     geom_label_repel(aes(label=Gene)) +
#     geom_abline(slope=1) +
#     xlab("Csn2 Correlation in Pregnancy") +
#     ylab("Csn2 Correlation in Tumorigenesis")
# 

dirtyPlot <- function(fplot){
    sub.gse <- quickGSE(fplot$Gene,fplot$fFDR)
    sub.gse$Fisher[sub.gse$Fisher=="< 1e-30"] <- "1e-30"
    sub.gse$Fisher <- -log10(as.numeric(sub.gse$Fisher))
    sub.gse %>%
	mutate(hitsPerc=(Significant/200)*100) %>% 
			    ggplot(aes(y=hitsPerc, 
				       x=Term, 
				       fill=Fisher
				       )) +
				geom_col() +
				expand_limits(x=0) +
				labs(y="Hits (%)", x="GO term", colour="p value", size="Count") +
				geom_text(aes(label=Term),y=1,
					  hjust=0,
					  color="grey30") +
				ylim(c(0,50)) +
				scale_fill_viridis(begin=1,end=0) +
				theme(axis.text.y=element_blank(),
				      axis.ticks.y=element_blank()) +
				coord_flip() 
}
fplot$Gene <- as.character(fplot$gene2)
fplot$fFDR <- ifelse(fplot$int2 & fplot$second_dist_to_med>0,
		     0.0001,1)
dirtyPlot(fplot)
```

## Enrichment of bottom left corner
```{r, message=FALSE,warning=FALSE, eval=FALSE}
fplot$Gene <- as.character(fplot$gene2)
fplot$fFDR <- ifelse(fplot$int2 & fplot$second_dist_to_med<0,
		     0.0001,1)
dirtyPlot(fplot)
```

### DE Analysis of high Csn2 cells 
```{r, fig.width=7, fig.height=7, warning=FALSE, message=FALSE}
cells <- cd$barcode[cd$Csn2Expr>5]
colnames(sce.full) <- sce.full$barcode
sce.sub <- sce.full[,cells]
sce.sub$Condition <- droplevels(sce.sub$Condition)
keep <- table(sce.sub$SampleID)>=30
keep <- names(keep[keep])
smrzd <- aggregateAcrossCells(sce.sub[,sce.sub$SampleID %in% keep],
			      id=colData(sce.sub[,sce.sub$SampleID %in% keep])[,c("SampleID")])
library(edgeR)
y <- DGEList(counts=counts(smrzd),
	     samples=colData(smrzd))
keep <- filterByExpr(y, group=y$samples$Experiment)
y <- y[keep,]
# Norm
y <- calcNormFactors(y)
mdl <- model.matrix(~ Experiment,y$samples)
y <- estimateDisp(y, mdl)

fit <- glmQLFit(y, mdl, robust=TRUE)
res <- glmQLFTest(fit, coef=ncol(mdl))
summary(decideTests(res, p.value=0.01))

top <- topTags(res,n=nrow(y))$table
top$Gene <- rownames(top)
library(ggrepel)

p2 <- ggplot(top, aes(x=logFC, y=-log10(FDR))) +
    geom_point(size=.5,color="grey80") +
    geom_point(data=top[1:50,], size=.9)+
    geom_label_repel(data=top[1:30,], aes(label=Gene)) +
    theme(text=element_text(size=12)) +
    xlab("logFC") +
    ylab("-log10(FDR)")


dirtyPlot <- function(fplot){
    sub.gse <- quickGSE(fplot$Gene,fplot$fFDR)
    sub.gse$Fisher[sub.gse$Fisher=="< 1e-30"] <- "1e-30"
    sub.gse$Fisher <- -log10(as.numeric(sub.gse$Fisher))
    sub.gse %>%
	mutate(hitsPerc=(Significant/200)*100) %>% 
			    ggplot(aes(y=hitsPerc, 
				       x=Term, 
				       fill=Fisher
				       )) +
				geom_col() +
				expand_limits(x=0) +
				labs(y="Hits (%)", x="GO term", colour="p value", size="Count") +
				geom_text(aes(label=Term),y=1,
					  hjust=0,
					  color="grey30") +
				ylim(c(0,80)) +
				scale_fill_viridis(begin=1,end=0) +
				theme(axis.text.y=element_blank(),
				      axis.ticks.y=element_blank()) +
				coord_flip() 
}

top$fFDR <- ifelse(top$FDR<0.01 & top$logFC>3,0.001,1)
p0 <- dirtyPlot(top) + ggtitle("Up in Tum")
top$fFDR <- ifelse(top$FDR<0.01 & top$logFC< -2,0.001,1)
p1 <- dirtyPlot(top) + ggtitle("Up in WT")
plot_grid(p2,p0,p1)
```
