---
title: "Figure 1"
author: "Karsten Bach"
date: '`r Sys.Date()`'
output:
  html_notebook:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: united
    highlight: tango
    code_folding: hide
---
***

# Main messages

In this file I will produce all the figure files for F1.

__Experimental setup of Tumorigenesis experiment__

- There is a need to understand pre-cancerous changes
- Sampling animals from a TNBC model pre-tumor formation and performing scRNAseq

__Inferring biological time__

- Age does not capture disease progression
- Inference of tumor time and grouping into stages

__Compositional changes during tumor development__

- Facetted UMAP as Overview
- Possible Ribbonplot / Barplot
I will keep the code and the plots for all of this in this document.
Including some further discussions.

```{r, message=FALSE,warning=FALSE}
library(scran)
library(scater)
library(ggplot2)
library(cowplot)
library(dplyr)
library(edgeR)
library(Matrix)
library(ggrepel)
library(ggrastr)
library(viridis)
theme_set(theme_cowplot())

sce.full <- readRDS("../../data/Integrated//Robjects/SCE_combined_final.rds")

sce <- sce.full[,sce.full$Experiment=="Tumorigenesis" & sce.full$Condition!="WTOld"]
pD <- data.frame(colData(sce))

outdr <- "../../data/figures/Figure1/"
```


# UMAP
The nice thing about the binning is that we can now more easily display the data in the form of UMAPs which quite pretty.
```{r, message=FALSE,warning=FALSE,fig.width=9,fig.height=6}
library(viridis)
p0 <- ggplot(pD, aes(x=UMAP1, y=UMAP2)) +
    geom_point(data=pD[,c("UMAP1","UMAP2")], color="grey80",size=.1,alpha=.2) +
    geom_point(size=.1, aes(color=Groups)) +
    facet_wrap(~ptimeBin,nrow=2) +
    scale_color_manual(values=levels(pD$GroupColors)) +
    guides(colour=guide_legend(title.position="top", override.aes = list(size=3))) +
    theme_void() +
    theme(legend.position="bottom",
	  legend.direction="horizontal",
	  strip.text=element_text(size=14))
p0
```


### Barplot for overall compartments
```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=7}
sce.sub <- sce.full[,sce.full$Condition %in% c("1","2","3","4","5","WTYoung","WTOld")]
pD <- data.frame(colData(sce.sub))
pD.add <- data.frame(colData(sce.sub))
pD.add$Condition <- "All"
pD <- droplevels(rbind(pD,pD.add))

sumry <- group_by(pD, Condition, Groups) %>%
    summarize(n=n()) %>%
    mutate(frac=n/sum(n)) %>%
    ungroup()

comb <- table(sumry$Groups,sumry$Condition) 
missing.comb <- which(comb==0, arr.ind=TRUE)
if (nrow(missing.comb)>0) {
    add.missing <- data.frame("Groups"=rownames(missing.comb),
		      "Condition"=colnames(comb)[missing.comb[,"col"]],
		      "frac"=0,
		      "n"=0)
    sumry <- rbind(sumry,add.missing)
}

sumry$Groups <- factor(sumry$Groups, levels=c("TumorEpithel","Basal","Luminal",
					      "Myeloid","Lymphoid","Fibroblast",
					      "Stroma"))

sumry$Colors <- plyr::mapvalues(as.character(sumry$Groups),
				as.character(unique(pD$Groups)),
				as.character(unique(pD$GroupColors)))

sumry <- arrange(sumry,Groups)

barplt <- ggplot(sumry, aes(x=Condition, y=frac, fill=Groups)) +
    geom_bar(stat="identity",color="black") +
    scale_fill_manual(values=unique(sumry$Colors)) +
    xlab("Condition") +
    ylab("Fraction of Cells") 
ggsave(paste0(outdr,"Barplot_Compartments.svg"))
```

```{r, message=FALSE, warning=FALSE,fig.width=7,fig.height=7}
dpi <- 800

ggplot(pD, aes(x=UMAP1, y=UMAP2, fill=CellTypesFinal)) +
    geom_point_rast(size=1.8,pch=21,stroke=.2,color="black") +
    scale_fill_manual(values=levels(pD$Color)) +
    theme_void() +
    theme(legend.position="none")

ggsave(paste0(outdr,"CellTypesUMAP.png"),dpi=dpi)

# With Annotation
X <- model.matrix(~0+pD$CellTypesFinal)
colnames(X) <- levels(pD$CellTypesFinal)
X <- t(t(X)/colSums(X))
celltyp_ctrs <- data.frame(crossprod(X, as.matrix(pD[,c("UMAP1","UMAP2")])))
celltyp_ctrs$CellType <- rownames(celltyp_ctrs)

ggplot(pD, aes(x=UMAP1, y=UMAP2, color=CellTypesFinal)) +
    geom_point_rast(size=.3, dpi=dpi) +
    scale_color_manual(values=levels(pD$Color)) +
    geom_text_repel(data=celltyp_ctrs,aes(label=CellType,color=NULL)) +
    theme(legend.position="none")
ggsave(paste0(outdr,"CellTypesUMAP_WithLabels.png"),dpi=dpi)

```

# Split UMAPs into individual files
```{r, message=FALSE, warning=FALSE,fig.width=7,fig.height=7}
pltUMAP <- function(pD.sub,pD) {
    p <- ggplot(pD, aes(x=UMAP1, y=UMAP2)) +
	geom_point_rast(color="grey80",size=1.8,dpi=dpi) +
	geom_point_rast(data=pD.sub, aes(fill=Groups), size=1.8,dpi=900,pch=21,stroke=.2) +
	scale_fill_manual(values=levels(pD.sub$GroupColors)) +
	theme_void() +
	theme(legend.position="none")
    return(p)
}


pDWt <- data.frame(colData(sce.full[,sce.full$Condition %in% c("WTYoung","WTOld")]))

set.seed(42)
# Select each pD individually and downsample
pD1 <- pD[pD$Condition==1,]
pD1 <- pD1[sample(nrow(pD1),12000),]

pD2 <- pD[pD$Condition==2,]
pD2 <- pD2[sample(nrow(pD2),12000),]

pD3 <- pD[pD$Condition==3,]
pD3 <- pD3[sample(nrow(pD3),12000),]

pD4 <- pD[pD$Condition==4,]
pD4 <- pD4[sample(nrow(pD4),12000),]

pD5 <- pD[pD$Condition==5,]

pDWtO <- pDWt[pDWt$Condition == "WTOld",]
pDWtO <- pDWtO[sample(nrow(pDWtO),12000),]
pDWtY <- pDWt[pDWt$Condition == "WTYoung",]
pDWtY <- pDWtY[sample(nrow(pDWtY),12000),]


ggsave(paste0(outdr,"UMAP_AllTumorigenesis.png"),pltUMAP(pD,pD),dpi=dpi)
ggsave(paste0(outdr,"UMAP_Bin_1.png"),pltUMAP(pD1,pD),dpi=dpi)
ggsave(paste0(outdr,"UMAP_Bin_2.png"),pltUMAP(pD2,pD),dpi=dpi)
ggsave(paste0(outdr,"UMAP_Bin_3.png"),pltUMAP(pD3,pD),dpi=dpi)
ggsave(paste0(outdr,"UMAP_Bin_4.png"),pltUMAP(pD4,pD),dpi=dpi)
ggsave(paste0(outdr,"UMAP_Bin_5.png"),pltUMAP(pD5,pD),dpi=dpi)
ggsave(paste0(outdr,"UMAP_Bin_WT_Young.png"),pltUMAP(pDWtY,pD),dpi=dpi)
ggsave(paste0(outdr,"UMAP_Bin_WT_Old.png"),pltUMAP(pDWtO,pD),dpi=dpi)

# Save one for later figure w all cells
pDall <- data.frame(colData(sce.full))

p <- ggplot(pDall, aes(x=UMAP1, y=UMAP2)) +
    geom_point_rast(aes(color=Groups), size=.3,dpi=900) +
    scale_color_manual(values=levels(pDall$GroupColors)) +
    theme_void() +
    theme(legend.position="none")

ggsave(paste0(outdr,"UMAP_TumorAndPregnancyCellGroups.png"), p, dpi=dpi)

library(RColorBrewer)
col.t <- brewer.pal(n=6,"PuRd")[5]
col.p <- brewer.pal(n=5,"GnBu")[5]

p <- ggplot(pDall, aes(x=UMAP1, y=UMAP2)) +
    geom_point_rast(aes(fill=Experiment), size=1.8,dpi=900,pch=21,stroke=.2) +
    scale_fill_manual(values=c(col.p,col.t)) +
    theme_void() +
    theme(legend.position="none")

ggsave(paste0(outdr,"UMAP_TumorAndPregnancyExperiment.png"), p, dpi=dpi)
```
