---
title: "Comunet on CellPhoneDB results"
author: "Karsten Bach"
date: '`r Sys.Date()`'
output:
  html_notebook:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: united
    highlight: tango
    code_folding: hide
---
***

# Read in Data
- First I have to load all the 10 (!) different conditions and convert to comunet objects
```{r, message=FALSE,warning=FALSE}
library(ggplot2)
library(COMUNET)
library(eulerr)
options(stringsAsFactors = F)
fldr <- "../../data/Downstream/CellPhoneDB_ImmEpi_out/"
conds <- c("WTYoung","4dG","9dG","14dG","WTOld","bin1","bin2","bin3","bin4")
complex_input <- read.csv("complex_input.csv")
complex_input$complex_name <- gsub("_"," ",complex_input$complex_name)
gene_input <- read.csv("gene_input.csv")

i.list <- list()
# Read in CPD
for (cond in conds) {
    dr <- paste0(fldr,cond)
    cpd <- read.csv(paste0(dr,"/significant_means.txt"),sep="\t",check.names=FALSE)
    cpd <- cpd[!duplicated(cpd$interacting_pair),]

    # add row names (interacting pairs)
    rownames(cpd) <- cpd$interacting_pair

    # transform 'receptor_a' colomn into boolean
    cpd$receptor_a  <- sapply(cpd$receptor_a
					     ,function(i){
					       if(i == "True") T else F
					     }
    )

    # transform 'receptor_b' colomn into boolean
    cpd$receptor_b  <- sapply(cpd$receptor_b
					     ,function(i){
					       if(i == "True") T else F
					     }
    )

    # Cond1 (AML328_d0)
    i.list[[cond]] <- convert_CellPhoneDB_output(CellPhoneDB_output = cpd,
					       complex_input = complex_input,
					       gene_input = gene_input)
}


plotEuler <- function(cond1, cond2, i.list=i.list) {
    library(eulerr)
    my_list <- list(i.list[[cond1]]$ligand_receptor_pair_df$pair
		    ,i.list[[cond2]]$ligand_receptor_pair_df$pair)
    names(my_list) <- c(cond1,
			cond2)
    euler_plot <- plot(euler(my_list)
		      ,quantities = list(fontsize = 18)
		      ,counts = TRUE
		      ,labels = list(fontsize = 18)
		     )
    print(euler_plot)
}

comp <- function(cond1, cond2) {
    # Run comparative analysis
    result <- comparative_analysis(cond1_weight_array = i.list[[cond1]]$weight_array
				  ,cond2_weight_array = i.list[[cond2]]$weight_array
				  ,cond1_ligand_receptor_pair_df = i.list[[cond1]]$ligand_receptor_pair_df
				  ,cond2_ligand_receptor_pair_df = i.list[[cond2]]$ligand_receptor_pair_df
				  ,cond1_nodes = i.list[[cond1]]$nodes
				  ,cond2_nodes = i.list[[cond2]]$nodes
				  ,cond1_name = cond1
				  ,cond2_name = cond2
				  )
    return(result)
}

getDiffPairs <- function(result,i.list=i.list) {
    # most different ligand-receptor pair
    out <- result$sorted_LRP_df$pair[result$sorted_LRP_df$presence == "shared" &
					   result$sorted_LRP_df$dissimilarity == max(result$sorted_LRP_df$dissimilarity)]
    return(out)
}


plotComGraphs <- function(i.list=i.list,LRP,m.frow=c(2,5),conds=NULL) {
    par(mfrow=m.frow)
    if (is.null(conds)) {
	conds <- names(i.list)
    }
    for (cond in conds) {
	plot_communication_graph(LRP=LRP,
				 weight_array=i.list[[cond]]$weight_array,
				 ligand_receptor_pair_df=i.list[[cond]]$ligand_receptor_pair_df,
				 nodes= i.list[[cond]]$node,
				 title=LRP,
				 subtitle=cond)
    }
}
```

# Intersections
We can first look at what interactions are present in the entire dataset per condition and see how they overlap with one another. Note that this does not take into account between which cell types the interaction takes place
```{r, message=FALSE,warning=FALSE,fig.width=14,fig.height=6}
library(UpSetR)
listInput <- lapply(i.list, function(x) x$ligand_receptor_pair_df$pair)
upset(fromList(listInput),nsets=length(listInput),order.by="freq",nintersects=60)
```

### Some hand picked ones
#### Potentially Interesing w regards to epithelium
```{r, message=FALSE,warning=FALSE,fig.width=20,fig.height=5}
LRP <- "TNFSF11:TNFRSF11A"
plotComGraphs(i.list=i.list,LRP=LRP)
LRP <- "IGF2:IGF2R"
plotComGraphs(i.list=i.list,LRP=LRP)
LRP <- "TGFB2:TGFbeta receptor1"
plotComGraphs(i.list=i.list,LRP=LRP)
LRP <- "LIF:LIFR"
plotComGraphs(i.list=i.list,LRP=LRP)
LRP <- "FGF1:FGFR2"
plotComGraphs(i.list=i.list,LRP=LRP)
```

#### Potentially Interesing w regards to immune
```{r, message=FALSE,warning=FALSE,fig.width=20,fig.height=5}
LRP <- "CDH1:KLRG1"
plotComGraphs(i.list=i.list,LRP=LRP)
LRP <- "CD70:CD27"
plotComGraphs(i.list=i.list,LRP=LRP)
LRP <- "IFNB1:Type I IFNR"
plotComGraphs(i.list=i.list,LRP=LRP)
LRP <- "IL10:IL10 receptor"
plotComGraphs(i.list=i.list,LRP=LRP)
LRP <- "CTLA4:CD80"
plotComGraphs(i.list=i.list,LRP=LRP)
LRP <- "XCL1:XCR1"
plotComGraphs(i.list=i.list,LRP=LRP)
```

# Differences in interactions
As they are now stored as graphs we can compute dissimilarity measures between the graphs and find interactions that are highly dissimilar (in a pair-wise fashion).
For example these are "very different" interactions in Bin1 vs. WTOld
```{r, message=FALSE,warning=FALSE,fig.width=20,fig.height=5}
res <- comp("WTOld","bin1")
diffs <- getDiffPairs(res)
```

### Comparisons

- This is from bin1 vs. WT1
```{r, message=FALSE,warning=FALSE,fig.width=20,fig.height=5}
for (dif in diffs) {
    plotComGraphs(i.list=i.list,LRP=dif)
}
plotComGraphs(i.list=i.list,LRP="SPP1:CD44")
plotComGraphs(i.list=i.list,LRP="SPP1:PTGER4")
plotComGraphs(i.list=i.list,LRP="SPP1:a4b1 complex")
plotComGraphs(i.list=i.list,LRP=diffs[5])
```

- This is bin1 vs. bin4
- cd96-nectin1
```{r, message=FALSE,warning=FALSE,fig.width=20,fig.height=5}
res <- comp("bin1","bin4")
diffs <- getDiffPairs(res)
for (dif in diffs) {
    plotComGraphs(i.list=i.list,LRP=diffs[3])
}
```
# Counting interactions
- Always been convinced that this doesn't make much sense.
```{r, message=FALSE,warning=FALSE}
sumlist <- list()
for (cond in conds) {
    i1 <- i.list[[cond]]
    intlist <- plyr::alply(i1$weight_array,3)
    intlist <- lapply(intlist, function(x) (x>0)*1)
    sums <- Reduce("+",intlist)
    sums <- sums + t(sums)
    diag(sums) <- diag(sums) / 2
    sumlist[[cond]] <- sums
}
# igr <- igraph::graph.adjacency(sums, mode="undirected", weighted=TRUE)
# ews <- 2.5
# ew <- (E(igr)$weight/max(E(igr)$weight))*ews
# igraph::plot.igraph(igr, layout=layout.circle(igr), edge.curved=TRUE,
#                     edge.width=ew)
library(pheatmap)
pheatmap(sumlist[["bin4"]]-sumlist[["bin1"]],
	 #          cluster_cols=FALSE,
	 #      WTOld    cluster_rows=FALSE,
	 clustering_method="ward.D2",
	 display_numbers=TRUE,
	 number_format="%.0f")
```

```{r, message=FALSE,warning=FALSE}
sumlist <- list()
for (cond in conds) {
    i1 <- i.list[[cond]]
    intlist <- plyr::alply(i1$weight_array,3)
    intlist <- lapply(intlist, function(x) (x>0)*1)
    sums <- Reduce("+",intlist)
    sums <- sums + t(sums)
    diag(sums) <- diag(sums) / 2
    sumlist[[cond]] <- sums
}
# igr <- igraph::graph.adjacency(sums, mode="undirected", weighted=TRUE)
# ews <- 2.5
# ew <- (E(igr)$weight/max(E(igr)$weight))*ews
# igraph::plot.igraph(igr, layout=layout.circle(igr), edge.curved=TRUE,
#                     edge.width=ew)
library(pheatmap)
pheatmap(sumlist[["9dG"]]-sumlist[["4dG"]],
	 #          cluster_cols=FALSE,
	 #      WTOld    cluster_rows=FALSE,
	 clustering_method="ward.D2",
	 display_numbers=TRUE,
	 number_format="%.0f")
```

```{r, message=FALSE,warning=FALSE}
i1 <- i.list[["bin1"]]
intlist1 <- plyr::alply(i1$weight_array,3)
intlist1 <- lapply(intlist1, function(x) (x>0)*1)
names(intlist1) <- i1$ligand_receptor_pair_df$pair

i4 <- i.list[["bin4"]]
intlist4 <- plyr::alply(i4$weight_array,3)
intlist4 <- lapply(intlist4, function(x) (x>0)*1)
names(intlist4) <- i4$ligand_receptor_pair_df$pair

comn <- intersect(names(intlist1),names(intlist4))
only1 <- setdiff(names(intlist1),names(intlist4))
only4 <- setdiff(names(intlist4),names(intlist1))

# Interaction-wise substraction of 4 - 1 for common interactions
comdf <- lapply(comn, function(LRP) intlist4[[LRP]]-intlist1[[LRP]])
sums <- Reduce("+",comdf)
sums <- sums + t(sums)
diag(sums) <- diag(sums) / 2

# For not shared interactions add interactions that are unique to 4 and substract interactions uq to 1
sum1 <- Reduce("+",intlist1[only1])
sum1 <- sum1 + t(sum1)
diag(sum1) <- diag(sum1) / 2

sum4 <- Reduce("+",intlist4[only4])
sum4 <- sum4 + t(sum4)
diag(sum4) <- diag(sum4) / 2

sumtot <- sums-sum1+sum4
source("../functions.R")
rownames(sumtot) <- renameForPlot(rownames(sumtot))
colnames(sumtot) <- renameForPlot(colnames(sumtot))
cairo_pdf("../../data/figures/Figure4/ChangeInInteractions.pdf")
pheatmap(sumtot,
	 #          cluster_cols=FALSE,
	 #      WTOld    cluster_rows=FALSE,
	 clustering_method="ward.D2",
	 display_numbers=TRUE,
	 number_format="%.0f",
	 legend=FALSE)
dev.off()


int <- lapply(comn, function(LRP) {
		    dff <- intlist4[[LRP]]-intlist1[[LRP]]
		    out <- dff["Lp","Avd"]>0 | dff["Avd","Lp"]>0
		    out
	 }
		    )
names(int) <- comn
int <- unlist(int)
int <- int[int]
plotComGraphs(i.list=i.list,LRP=names(int)[26])

int <- lapply(comn, function(LRP) {
		    dff <- intlist4[[LRP]]-intlist1[[LRP]]
		    out <- dff["Hs","Avd"]<0 | dff["Avd","Hs"]<0
		    out
	 }
		    )
names(int) <- comn
int <- unlist(int)
int <- int[int]
plotComGraphs(i.list=i.list,LRP=names(int)[8])
plotComGraphs(i.list=i.list,LRP="TNFSF11:TNFRSF11A")
plotComGraphs(i.list=i.list,LRP="TNFSF4:TNFRSF4")
plotComGraphs(i.list=i.list,LRP="SPP1:CD44")

# For the ones that are not common
int <- lapply(only4, function(LRP) {
		    dff <- intlist4[[LRP]]
		    out <- dff["Mo3","Avd"]>0 | dff["Avd","Mo3"]>0
		    out
	 }
		    )
names(int) <- only4
int <- unlist(int)
int <- int[int]
plotComGraphs(i.list=i.list,LRP=names(int)[1])

plotComGraphs(i.list=i.list["bin1"],LRP="SPP1:CD44")

#Get in the colors
library(scran)
pD <- data.frame(colData(readRDS("../../data/Integrated/Robjects/SCE_combined_celltypes.rds")))
cols <- levels(pD$Colors)
names(cols) <- levels(pD$CellTypesFinal)

altPlot <- function(i.list,LRP,cond, onlyEpithelium=FALSE,colscheme=cols) {
    require(ggraph)
    lrp <- LRP
    il <- i.list[[cond]]
    # Trying alternative plots
    wa <- il$weight_array
    lr <- il$ligand_receptor_pair_df
    m <- wa[,,lrp]
    dm <- COMUNET:::calc_degrees(weight_array=m)
    dm <- COMUNET:::array2matrix(dm)
    ordr <- levels(factor(colnames(m)))
    ordr[ordr=="Basal"] <- "Bsl1"
    colnames(m) <- renameForPlot(colnames(m))
    rownames(m) <- renameForPlot(rownames(m))
    if (onlyEpithelium) {
	m <- m[c("Avd","Lp","Basal","Hs"),c("Avd","Lp","Basal","Hs")]
    }
    igr <- igraph::graph.adjacency(m, mode="directed", weighted=TRUE)
    # Width
    edge.width <- (E(igr)$weight/max(E(igr)$weight)) * 2.5
    ggraph(igr, layout="linear", circular=TRUE) +
	geom_edge_arc(aes(start_cap = label_rect(node1.name), end_cap = label_rect(node2.name),
			  width=weight),
		       arrow = arrow(type = "closed", length = unit(3, 'mm')),
		       color="grey50",
		       strength=5) +
	geom_edge_loop(aes(start_cap = label_rect(node1.name),
			   end_cap = label_rect(node2.name),
			   span=90, direction=90, strength=.3),color="grey50",
		       arrow = arrow(type = "closed", length = unit(3, 'mm'))) +
	geom_node_point(size=9,pch=21,aes(fill=name),color="black",stroke=1) +
	geom_node_text(aes(label = name), 
		       size = 9, 
		       vjust = 1.3) +
	theme_void() +
        coord_cartesian(xlim=c(-1.2,1.3), ylim=c(-1.2,1.2)) +
	scale_edge_width_continuous(range=c(.5,2)) +
	scale_fill_manual(values=unname(colscheme[ordr])) +
	ggtitle(paste0(LRP," in ",cond)) +
       guides(edge_width = FALSE,
	      fill=FALSE) 
}

altPlot(i.list,"SPP1:CD44","bin1")
ggsave("../../data/figures/Figure4/Spp1_bin1.pdf")
altPlot(i.list,"SPP1:CD44","WTOld")
ggsave("../../data/figures/Figure4/Spp1_WTOld.pdf")

altPlot(i.list,"TNFSF11:TNFRSF11A","9dG")
ggsave("../../data/figures/Figure4/Rank_9dG.pdf")
altPlot(i.list,"TNFSF11:TNFRSF11A","bin4")
ggsave("../../data/figures/Figure4/Rank_bin4.pdf")

altPlot(i.list,"IGF2:IGF2R","bin4")
ggsave("../../data/figures/Figure4/Igf2_bin4.pdf")
altPlot(i.list,"IGF2:IGF2R","9dG")
ggsave("../../data/figures/Figure4/Igf2_9dG.pdf")

# On special request by the boss
altPlot(i.list,"LIF:LIFR","bin1")
ggsave("../../data/figures/Figure4/Lif_bin1.pdf",width=9,height=9)

altPlot(i.list,"LIF:LIFR","bin4")
ggsave("../../data/figures/Figure4/Lif_bin4.pdf",width=9,height=9)

altPlot(i.list,"FGF1:FGFR1","bin1")
ggsave("../../data/figures/Figure4/Fgf1_bin1.pdf",width=9,height=9)

altPlot(i.list,"FGF1:FGFR1","bin4")
ggsave("../../data/figures/Figure4/Fgf1_bin4.pdf",width=9,height=9)

# alternative layouts
# ggraph(igr, layout="linear", circular=TRUE) +
#     geom_edge_arc(aes(alpha=..index..,
#                       width=weight),
#                    color="black",
#                    strength=5) +
#     geom_node_text(aes(label=name)) +
#     theme_graph() +
#     scale_edge_width_continuous(range=c(.5,2)) +
#     scale_edge_alpha('Edge direction', guide = 'edge_direction') +
#    guides(edge_width = FALSE) 
# 
# ggraph(igr, layout="circle") +
#     geom_edge_fan(aes(start_cap = label_rect(node1.name), end_cap = label_rect(node2.name),
#                       width=weight),
#                    arrow = arrow(type = "closed", length = unit(3, 'mm')),
#                    color="grey50",
#                    strength=5) +
#     geom_edge_loop() +
#     geom_node_text(aes(label=name)) +
#     theme_graph() +
#     scale_edge_width_continuous(range=c(.5,2)) +
#     scale_edge_color_manual(values=cols) +
#    guides(edge_width = FALSE) 

```

# Try different visualisations for sumtot
```{r, message=FALSE,warning=FALSE}
cellCellContactMap = function(pmat, order = NULL, exclude = NULL) {
    # this function takes output from cellCellContact() 
    # and gives a ggplot object of the graph
    # out = mat_p_sym
    # if order not given then perform hclust to get ordering
    require(reshape)
    require(ggplot2)
    mat_p_df = melt(pmat)
    colnames(mat_p_df) <- c("CellType_1","CellType_2", "NInteractions")
    if (is.null(order)) {
        hc = hclust(dist(pmat))
        mat_p_df$CellType_1 <- factor(mat_p_df$CellType_1, levels = 
                                            hc$labels[hc$order])
        mat_p_df$CellType_2 <- factor(mat_p_df$CellType_2, levels = 
                                            hc$labels[hc$order])
    } else {
        mat_p_df$CellType_1 <- factor(mat_p_df$CellType_1, levels = 
                                            order)
        mat_p_df$CellType_2 <- factor(mat_p_df$CellType_2, levels = 
                                            order)
    }
    mat_p_df$keep = as.numeric(mat_p_df$CellType_1) >= as.numeric(mat_p_df$CellType_2) 
    g = ggplot(subset(mat_p_df, keep & 
                          (!CellType_1 %in% exclude) & 
                          (!CellType_2 %in% exclude)),
               aes(x = CellType_1, y = CellType_2, fill =NInteractions)) + 
        geom_tile() +
       # geom_text(aes(label = Sig, y = CellType_2 - 0.025), size = 20) +
        # geom_point(pch = "*", size = 10, data = subset(mat_p_df, Sig == "*" & keep)) +
        theme_classic() +
        theme(axis.line = element_blank()) +
        theme(axis.ticks = element_blank()) +
        # scale_y_continuous(position = "right") +
        scale_y_discrete(position = "right") +
        theme(axis.text = element_text(size = 14)) +
        theme(axis.title = element_blank()) +
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
        coord_fixed() +
        # guides(fill = guide_legend(title = "")) +
        scale_fill_gradient2(high ="#9E0142",
			     mid = "white",
			     low = "#5182BB",
			    # na.value = "white",
                             midpoint = 0, #limits = c(0,1),
                            # breaks = c(0,1),
			    # labels = c("Segregated", "Integrated")
			     ) +
        theme(legend.position = "top") + 
        theme(legend.text = element_text(size = 15)) +
        theme(legend.key.width = unit(1, "in")) +
        guides(fill = guide_colourbar(title.position = "top",
                                      title = "",
                                      title.hjust = 0.5,
                                      ticks = FALSE,
                                      reverse = TRUE)) +
        NULL
    # print(g)
    return(g)
}

cellCellContactMap(sumtot)
ggsave("../../data/figures/Figure4/ChangeInInteractionsTriangle.pdf")
```
