---
title: "Testing for Changes in abundance"
author: "Karsten Bach"
date: '`r Sys.Date()`'
output:
  html_notebook:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: united
    highlight: tango
    code_folding: hide
---
***

# Load Data
```{r, message=FALSE,warning=FALSE}
library(scran)
library(scater)
library(ggplot2)
library(cowplot)
library(dplyr)
library(Matrix)
source("../functions.R")
theme_set(theme_cowplot())

sce <- readRDS("../../data/Integrated//Robjects/SCE_combined_final.rds")
sce <- sce[,sce$Experiment=="Tumorigenesis"]

sce$CellTypes <- sce$CellTypesFinal
outdr <- "../../data/figures/Figure2/"
```

# Differences in cluster abundance over tumor time
I have changed to modeling the cluster abundances with rlm instead of the general NB GLM from edgeR to protect against outliers at the exterme ends of the explanatory variable.
```{r, fig.width=7, fig.height=7, warning=FALSE, message=FALSE}
pD <- data.frame(colData(sce))
pD <- droplevels(pD[pD$Condition!="WTOld",])
sumry <- group_by(pD, Condition, CellTypes) %>%
    summarize(n=n()) %>%
    mutate(frac=n/sum(n)) %>%
    ungroup()

comb <- table(sumry$CellTypes,sumry$Condition) 
missing.comb <- which(comb==0, arr.ind=TRUE)
if (nrow(missing.comb)>0) {
    add.missing <- data.frame("CellTypes"=rownames(missing.comb),
		      "Condition"=colnames(comb)[missing.comb[,"col"]],
		      "frac"=0,
		      "n"=0)
    sumry <- rbind(sumry,add.missing)
}

sumry$Colors <- plyr::mapvalues(as.character(sumry$CellTypes),
				as.character(unique(pD$CellTypes)),
				as.character(unique(pD$Colors)))

combi <- unique(paste0(substr(pD$MajorGroups,1,3),"-",as.character(pD$CellTypes)))
names(combi) <- unlist(lapply(strsplit(combi,"-"),function(x) x[[2]]))
sumry$CellTypes <- plyr::mapvalues(as.character(sumry$CellTypes),
				   as.character(names(combi)),
				   as.character(combi))

sumry <- arrange(sumry, Condition, CellTypes) %>%
    group_by(Condition) %>%
    mutate(cumfrac=cumsum(frac)) %>%
    ungroup()

sumry$Condition <- factor(sumry$Condition,levels=c(5,4,3,2,1))
levels(sumry$Condition)[1] <- "T"
barplt <- ggplot(sumry, aes(x=Condition, y=frac*100, fill=CellTypes)) +
    geom_bar(stat="identity",color="black") +
    scale_fill_manual(values=unique(sumry$Colors)) +
    xlab("") +
    ylab("") +
    coord_flip() +
    theme_pub() 
#    theme(legend.position="none") 
barplt

ggsave(paste0(outdr,"Barplot_Tumor_all.pdf"))#,width=5,height=5)




pD <- droplevels(pD[pD$Condition!="5",])
cluster.counts <- table(pD$CellTypes, pD$SampleID)
cluster.counts <- cluster.counts[rowMeans(cluster.counts)>10,]
cluster.counts <- as(cluster.counts,"matrix")

library(dplyr)
sumry <- group_by(pD, SampleID, ptime) %>%
    summarize()

# EdgeR
library(edgeR)
# Set up DGE
y.ab <- DGEList(cluster.counts)
y.ab$samples$SampleID <- colnames(y.ab)
y.ab$samples <- left_join(y.ab$samples,sumry)


# I am rescaling ptime so that the coefficients of the model are in a more interpretable space, this does not change the results just the scale of the coeffs
y.ab$samples$ptime <- y.ab$samples$ptime/max(y.ab$samples$ptime)

tab <- data.frame()

for (ct in rownames(cluster.counts)) {
    library(MASS)
    library(sfsmisc)
    y.ab$samples$Counts <- cpm(y.ab, log=TRUE)[ct,]
    fit <- rlm(formula=Counts ~  ptime, data=y.ab$samples,maxit=100)
    ftest <- f.robftest(fit,var="ptime")
    tmp <- data.frame("CellTypes"=ct,
		      "lfc"=fit$coefficients[length(fit$coefficient)],
		      "intercept"=fit$coefficients[1],
		      "PVal"=ftest$p.value)
    tab <- rbind(tab,tmp)
}
tab$FDR <- p.adjust(tab$PVal, method="BH")
tab[order(tab$FDR,decreasing=FALSE),]

addCol <- data.frame("CellTypes"=levels(pD$CellTypesFinal),
		     "Colors"=levels(pD$Colors))
rownames(addCol) <- addCol$CellTypes
addCol <- addCol[as.character(tab$CellTypes),]
tab$Colors <- addCol$Colors
tab$Colors <- factor(tab$Colors, levels=tab$Colors)
tab$CellTypes <- factor(tab$CellTypes, levels=tab$CellTypes)
library(ggrepel)
tab$CellTypes <- renameForPlot(tab$CellTypes)
volc <- ggplot(tab, aes(x=lfc, y=-log10(FDR),fill=CellTypes)) +
    geom_point(pch=21,size=4) +
    geom_hline(yintercept=1,size=1.1, lty="dashed") +
    geom_text_repel(data=tab[tab$FDR<0.1,], label=tab[tab$FDR<0.1,"CellTypes"],aes(fill=NULL),
		     size=6, color="grey30",force=9) +
    scale_fill_manual(values=levels(tab$Colors)) +
    theme_pub() +
    theme(legend.position="none") +
    xlab("Log2(FC)")
volc

tab.tum <- tab
# Save for Shiny server
write.csv(tab,"../../Shiny/data/TumourDA.csv")
ggsave(paste0(outdr,"Volcano.pdf"),volc,width=7.6,height=6.6)
```

# Pregnancy
```{r, message=FALSE,warning=FALSE}
sce <- readRDS("../../data/Integrated/Robjects/SCE_combined_final.rds")
sce <- sce[,sce$Experiment=="Pregnancy"]
pD <- data.frame(colData(sce))
```

## Pregnancy robust
As I intend to compare pregnancy vs tumorigenesis I will try the same rlm approach as above but for the gestation timepoints.
The only difference here is that I normalize for compositional biases introduced by the strong changes in a couple of clusters, if this normalization is left out almost everything changes.

```{r, fig.width=7, fig.height=7, warning=FALSE, message=FALSE}
cluster.counts <- table(pD$CellTypes, pD$SampleID)
cluster.counts <- cluster.counts[rowMeans(cluster.counts)>10,]
cluster.counts <- as(cluster.counts,"matrix")

library(edgeR)
# Set up DGE
y.ab <- DGEList(cluster.counts)
y.ab$samples$SampleID <- rownames(y.ab$samples)
y.ab$samples$Condition <- unlist(lapply(strsplit(rownames(y.ab$samples),"_"),function(x) x[1]))
y.ab$samples$Condition <- gsub("CTRL","0",y.ab$samples$Condition)
y.ab$samples$Condition <- gsub("dG","",y.ab$samples$Condition)
y.ab$samples$Condition <- as.numeric(y.ab$samples$Condition)

y.ab <- calcNormFactors(y.ab)
tab <- data.frame()
for (ct in rownames(cluster.counts)) {
    library(MASS)
    library(sfsmisc)
    y.ab$samples$Counts <- cpm(y.ab, log=TRUE)[ct,]
    fit <- rlm(formula=Counts ~  Condition, data=y.ab$samples,maxit=100)
    ftest <- f.robftest(fit,var="Condition")
    tmp <- data.frame("CellTypes"=ct,
		      "lfc"=fit$coefficients[length(fit$coefficient)],
		      "intercept"=fit$coefficients[1],
		      "PVal"=ftest$p.value)
    tab <- rbind(tab,tmp)
}

tab$FDR <- p.adjust(tab$PVal, method="BH")
tab[order(tab$FDR,decreasing=FALSE),]

addCol <- data.frame("CellTypes"=levels(pD$CellTypes),
		     "Colors"=levels(pD$Colors))
rownames(addCol) <- addCol$CellTypes
addCol <- addCol[tab$CellTypes,]
tab$Colors <- addCol$Colors
tab$Colors <- factor(tab$Colors, levels=tab$Colors)
tab$CellTypes <- factor(tab$CellTypes, levels=tab$CellTypes)
library(ggrepel)
volc <- ggplot(tab, aes(x=lfc, y=-log10(FDR),fill=CellTypes)) +
    geom_point(pch=21,size=4) +
    geom_hline(yintercept=1) +
    geom_label_repel(data=tab[tab$FDR<0.1,], label=tab[tab$FDR<0.1,"CellTypes"],aes(fill=NULL)) +
    scale_fill_manual(values=levels(tab$Colors)) +
    theme(legend.position="none") 
volc

tab.preg <- tab
```

## Tumor vs Pregnancy
```{r, fig.width=7, fig.height=7, warning=FALSE, message=FALSE}
comp <- tab.preg[,c("CellTypes","Colors","lfc","FDR")]
colnames(comp)[3-4] <- paste0(colnames(comp)[3-4],".Pregnancy")
add <- tab.tum[,c("CellTypes","lfc","FDR")]
colnames(add)[2-3] <- paste0(colnames(add)[2-3],".Tumor")
comp <- dplyr::full_join(comp,add)
comp$significant <- "None"
comp$significant[comp$FDR.Tumor < 0.1] <- "Tumorigenesis"
comp$significant[comp$FDR.Pregnancy < 0.1] <- "Pregnancy"
comp$significant[comp$FDR.Pregnancy < 0.1 & comp$FDR.Tumor < 0.1] <- "Both"

comp$lfc.Tumor[is.na(comp$lfc.Tumor)] <- 0
comp$FDR.Tumor[is.na(comp$FDR.Tumor)] <- 1
comp$lfc.Pregnancy[is.na(comp$lfc.Pregnancy)] <- 0
comp$FDR.Pregnancy[is.na(comp$FDR.Pregnancy)] <- 1

comp$lfc.Tumor[comp$lfc.Tumor>0] <- (comp$lfc.Tumor[comp$lfc.Tumor>0])/max(comp$lfc.Tumor[comp$lfc.Tumor>0])
comp$lfc.Tumor[comp$lfc.Tumor<0] <- -(-comp$lfc.Tumor[comp$lfc.Tumor<0])/max(-comp$lfc.Tumor[comp$lfc.Tumor<0])
comp$lfc.Pregnancy[comp$lfc.Pregnancy>0] <- (comp$lfc.Pregnancy[comp$lfc.Pregnancy>0])/max(comp$lfc.Pregnancy[comp$lfc.Pregnancy>0])
comp$lfc.Pregnancy[comp$lfc.Pregnancy<0] <- -(-comp$lfc.Pregnancy[comp$lfc.Pregnancy<0])/max(-comp$lfc.Pregnancy[comp$lfc.Pregnancy<0])
# 
# comp$lfc.Tumor <- scale(comp$lfc.Tumor)
# comp$lfc.Pregnancy <- scale(comp$lfc.Pregnancy)


comp$significant <- factor(comp$significant,levels=c("None","Both","Pregnancy","Tumorigenesis"))

ggplot(comp, aes(x=lfc.Pregnancy, y=lfc.Tumor, color=significant)) +
    geom_point(size=2) + 
    geom_label_repel(data=comp[comp$significant!="None",],aes(label=CellTypes)) +
    geom_abline(slope=1,intercept=0) +
    geom_vline(xintercept=0,lty="dashed") +
    geom_hline(yintercept=0,lty="dashed") +
    xlab("Scaled LFC during Pregnancy") +
    ylab("Scaled LFC during Tumorigenesis") +
    scale_color_manual(values=c("grey80","black","dodgerblue","red")) +
    coord_fixed(1)
```
