---
title: "Figure 1"
author: "Karsten Bach"
date: '`r Sys.Date()`'
output:
  html_notebook:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: united
    highlight: tango
    code_folding: hide
---
***

# Main messages

In this file I will produce all the figure files for F1.

__Experimental setup of Tumorigenesis experiment__

- There is a need to understand pre-cancerous changes
- Sampling animals from a TNBC model pre-tumor formation and performing scRNAseq

__Inferring biological time__

- Age does not capture disease progression
- Inference of tumor time and grouping into stages

__Compositional changes during tumor development__

- Facetted UMAP as Overview
- Possible Ribbonplot / Barplot
I will keep the code and the plots for all of this in this document.
Including some further discussions.

```{r, message=FALSE,warning=FALSE}
library(scran)
library(scater)
library(ggplot2)
library(cowplot)
library(dplyr)
library(edgeR)
library(Matrix)
library(ggrepel)
library(ggrastr)
library(viridis)
theme_set(theme_cowplot())

sce.full <- readRDS("../../data/Integrated//Robjects/SCE_combined_final.rds")
sce.full$CellTypes <- factor(sce.full$CellTypesFinal)
sce.full$Colors <- factor(sce.full$Colors)
sce.full$Batch <- as.character(sce.full$Batch)
sce.full$Groups <- factor(sce.full$Groups)

sce <- sce.full[,sce.full$Experiment=="Tumorigenesis"]
ptime <- read.csv("../../data/Integrated/Robjects/TumorTime.csv")[,-1]

sce <- sce[,sce$Condition %in% ptime$Condition]
pD <- data.frame(colData(sce))
pD <- dplyr::left_join(pD,ptime)
colData(sce) <- DataFrame(pD)
getPC <- function(sce, getPCObject=FALSE) {
    sce.sum <- aggregateAcrossCells(sce,
			  id=DataFrame(Condition=sce$Condition),
			  use_exprs_values="counts",
			  average=FALSE)

    sce.counts <- counts(sce.sum)
    y.sce <- DGEList(counts=sce.counts, samples=data.frame(colData(sce.sum)))
    keep <- filterByExpr(y.sce)
    y.sce <- y.sce[keep,]
    y.sce <- calcNormFactors(y.sce)

    logcpm <- cpm(y.sce, log=TRUE)
    # 
    pD <- data.frame(colData(sce.sum))
    #
    pcs <- prcomp(t(logcpm))
    pca <- data.frame(pcs$x)
    pca[,"Condition"] <- pD$Condition
    pca$Age <- pD$Age
    pca$Batch <- pD$Batch
    # pca$CellTypes <- y$samples$CellTypes
    rownames(pca) <- pca$Condition
    if (getPCObject) {
	return(pcs)
    } else {
	return(pca)
    }
}
```


# Pseudo tumor time
In the first figure we would need to define the tumor time that we use in any subsequent analysis.
Here this is done on the pseudo-bulk samples across all cells per sample.
I would show this definitely as PCA, and potentially also as this "timeline".

```{r, message=FALSE,warning=FALSE,fig.width=6,fig.height=3}
sumry <- group_by(pD, Batch, Age, Condition) %>%
    summarize(n=n())

sumry <- sumry[grepl("WKBR",sumry$Condition),]
sumry <- arrange(sumry,Age)
sumry$Condition <- factor(sumry$Condition,
			  levels=sumry$Condition)
cols <- data.frame(Age=unique(sumry$Age))
cols$cols <- plasma(nrow(cols),begin=1,end=0)

sumry <- left_join(sumry,cols)

```

## PCA to construct tumor time
- PCA is computed on all cells per sample
```{r, message=FALSE,warning=FALSE,fig.width=7,fig.height=7}
pcs.all <- getPC(sce)
pcs.all$PC1 <- -pcs.all$PC1
pcs.all$Condition <- rownames(pcs.all)

pcs.all <- left_join(pcs.all,sumry[,c("Condition","cols")])
pcs.all <- arrange(pcs.all,Age)
# pcs.all$Age <- factor(pcs.all$Age,levels=c("30","33","38","41","46","48"))
colrd <- pcs.all$cols
names(colrd) <- pcs.all$Age
pcs.all$Tumor <- grepl("TM",pcs.all$Condition)
pcs.all$Tumor <- ifelse(pcs.all$Tumor,"Tumor","Mammary Gland")

p1 <- ggplot(pcs.all, aes(x=PC1, y=PC2,fill=Tumor)) +
    geom_point(size=4,pch=21) +
    scale_fill_manual(values=c("grey80","grey30")) +
    guides(fill = guide_legend(title = "")) +
    coord_fixed(1)
p1
```

## Timeline
```{r, message=FALSE,warning=FALSE,fig.width=6,fig.height=3}
timeplt_byCond <- ggplot(sumry, aes(x=as.numeric(Age), y=0)) +
    geom_segment(aes(x=25, y=0, xend=55, yend=0,
		     color=NULL),
		 arrow=arrow(length=unit(0.2, "cm"))) +
    geom_dotplot(aes(fill=Condition, group=Condition), binwidth=1, 
	     stackgroups=TRUE, binpositions="all") +
    xlim(25,55) +
    ylim(0,6) +
    xlab("Age") +
    scale_fill_manual(values=sumry$cols) +
    theme(axis.title.y=element_blank(),
	  axis.line.x=element_blank(),
	  axis.line.y=element_blank(),
	  axis.text.y=element_blank(),
	  legend.position="none",
	  axis.ticks.y=element_blank(),
	  axis.ticks.x=element_line(size=1, linetype="solid"),
	  axis.ticks.length.x=unit(.5,"cm")) +
    coord_fixed(1)

pcs.all$ptime <- (pcs.all$PC1-min(pcs.all$PC1))/max(pcs.all$PC1-min(pcs.all$PC1)) * 100

sumry <- left_join(sumry,pcs.all[,c("Condition","ptime")])
sumry$Condition <- factor(sumry$Condition,
			  levels=sumry$Condition)

pseudotimeplt_byCond <- ggplot(sumry, aes(x=ptime, y=0)) +
    geom_segment(aes(x=0, y=0, xend=110, yend=0,
		     color=NULL),
		 arrow=arrow(length=unit(0.2, "cm"))) +
    geom_dotplot(aes(fill=Condition, group=Condition), binwidth=1, 
		 stackgroups=TRUE, binpositions="all", dotsize=3) +
    xlim(0,110) +
    ylim(0,22) +
    xlab("Pseudotime") +
    scale_fill_manual(values=sumry$cols) +
    ggtitle("Inferred Order") +
    theme(axis.title.y=element_blank(),
	  axis.line.x=element_blank(),
	  axis.line.y=element_blank(),
	  axis.text.y=element_blank(),
	  legend.position="none",
	  axis.ticks.y=element_blank(),
	  axis.ticks.x=element_line(size=1, linetype="solid"),
	  axis.ticks.length.x=unit(.5,"cm")) +
    scale_x_continuous(breaks=c(0,25,50,75,100)) +
    coord_fixed(1)

timeplt_byCond
pseudotimeplt_byCond
p2 <- plot_grid(timeplt_byCond, pseudotimeplt_byCond,ncol=2)
```


## PC1 Loadings
Finally, we can directly look at the loadings of PC1 to see if the genes with high positive loadings can be related to the process of tumor formation.
In fact, we see cell division and innate immune response popping up.
The negative loadings are not so easy to interpret.
```{r, message=FALSE, warning=FALSE}
pcs <- getPC(sce, getPCObject=TRUE)


ld.pc1 <- pcs$rotation[,1]
ld.pc1.ord <- ld.pc1[order(ld.pc1,decreasing=FALSE)]
fp <- data.frame("Gene"=names(ld.pc1.ord),
		 "Loading"=ld.pc1.ord)
fp$Index <- c(1:nrow(fp))
library(ggrepel)
p.ld <- ggplot(fp, aes(x=Index, y=Loading)) +
    geom_point(color="grey80") +
    geom_point(data=fp[abs(fp$Loading)>0.02,],color="black") +
    geom_text_repel(data=fp[c(1:10,(nrow(fp)-10):nrow(fp)),], aes(label=Gene)) +
    geom_hline(yintercept=0.02,color="red",lty="dashed") +
    geom_hline(yintercept=-0.02,color="red",lty="dashed") +
    ylab("PC1 Loading") +
    coord_flip()

ps <- ld.pc1[ld.pc1 > 0.02]
ps <- ps[order(ps,decreasing=TRUE)]
neg <- ld.pc1[ld.pc1 < -0.02]
neg <- neg[order(neg)]

topLoads <- list("Positive Loading"=ps, "Negative Loading"=neg)# topGO
univrs <- names(ld.pc1)
out <- list()

for (i in seq_along(topLoads)) {
    gens <- names(topLoads[[i]])
    subst <- names(topLoads)[[i]]
    # Gene universe
    alG <- factor(as.numeric(univrs %in% gens))
    names(alG) <- univrs

    # ---- GOanalysis ----

    library(topGO)
    #prepare Data for topGO
    GO.data <- new("topGOdata", description="Lib GO",ontology="BP", allGenes=alG, 
		   annot=annFUN.org, mapping="org.Mm.eg.db",
		   nodeSize=20, ID="symbol")
    result.classic <- runTest(GO.data, statistic="fisher")
    res <- GenTable(GO.data, Fisher.classic=result.classic, orderBy="topgoFisher", topNodes=50, numChar=300)

    res$Fisher.classic <- as.numeric(res$Fisher.classic)
    res$Term <- factor(res$Term, levels=res$Term[order(res$Fisher.classic,decreasing=TRUE)])
    out[[subst]] <- res[1:15,]
}

p.ld
```


```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
p.ld.2 <- out[[2]] %>%  # this only works because they are ordered by fisher classic
			mutate(hitsPerc=Significant*100/length(gens)) %>% 
			ggplot(aes(y=hitsPerc, 
				   x=Term, 
				   fill=Fisher.classic
				   )) +
			    geom_col() +
			    expand_limits(x=0) +
			    labs(y="Hits (%)", x="GO term", colour="p value", size="Count") +
			    geom_text(aes(label=Term),y=1,
				      hjust=0,
				      color="grey30") +
			    ylim(c(0,20)) +
			    scale_fill_viridis(begin=1,end=0) +
			    theme(axis.text.y=element_blank(),
				  axis.ticks.y=element_blank()) +
			    coord_flip() 
			
p.ld.2
```




## Binning
Here we bin the data along PC1 (or tumortime) into 4 groups pre-tumor formation and keep the two tumors as two separate samples.
```{r, message=FALSE,warning=FALSE,fig.width=7,fig.height=7}
binbounds <- levels(cut(ptime$ptime[ptime$ptime<80],4))
binbounds <- unlist(lapply(strsplit(binbounds,","),function(x) x[2]))
binbounds <- as.numeric(gsub("]","",binbounds))
binbounds <- c(min(ptime$ptime),binbounds)
pseudotimeplt_byCond + geom_vline(xintercept=binbounds, lty="dashed")

binbounds <- levels(cut(pcs.all$PC1[pcs.all$PC1<50],4))
binbounds <- unlist(lapply(strsplit(binbounds,","),function(x) x[2]))
binbounds <- as.numeric(gsub("]","",binbounds))
binbounds <- c(min(pcs.all$PC1),binbounds)

p1_col <- p1 +
    geom_rect(aes(xmin=-Inf,
		      xmax = binbounds[2],
		      ymin = -Inf,
		      ymax = Inf), fill = 'pink', alpha = 0.02) +
    geom_rect(aes(xmin=binbounds[2],
		      xmax = binbounds[3],
		      ymin = -Inf,
		      ymax = Inf), fill = 'pink', alpha = 0.06) +
    geom_rect(aes(xmin=binbounds[3],
		      xmax = binbounds[4],
		      ymin = -Inf,
		      ymax = Inf), fill = 'pink', alpha = 0.10) +
    geom_rect(aes(xmin=binbounds[4],
		      xmax = binbounds[5],
		      ymin = -Inf,
		      ymax = Inf), fill = 'pink', alpha = 0.20) +
    geom_rect(aes(xmin=binbounds[5],
		      xmax = Inf,
		      ymin = -Inf,
		      ymax = Inf), fill = 'red', alpha = 0.03) +
    geom_vline(xintercept=binbounds, lty="dashed") +
    geom_point(size=4,pch=21) #This is just laziness
p1_col
```

# UMAP
The nice thing about the binning is that we can now more easily display the data in the form of UMAPs which quite pretty.
```{r, message=FALSE,warning=FALSE,fig.width=9,fig.height=6}

pD$GroupColors <- plyr::mapvalues(pD$Groups,
				levels(pD$Groups),
				c("#00BFFF","#37A055","#3787C0","#F78757","#FF4E03",
				  "#737373","#800080"))
library(viridis)
p0 <- ggplot(pD, aes(x=UMAP1, y=UMAP2)) +
    geom_point(data=pD[,c("UMAP1","UMAP2")], color="grey80",size=.1,alpha=.2) +
    geom_point(size=.1, aes(color=Groups)) +
    facet_wrap(~ptimeBin,nrow=2) +
    scale_color_manual(values=levels(pD$GroupColors)) +
    guides(colour=guide_legend(title.position="top", override.aes = list(size=3))) +
    theme_void() +
    theme(legend.position="bottom",
	  legend.direction="horizontal",
	  strip.text=element_text(size=14))
p0
```


```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=7}
library(dplyr)
# Ribbon plot
pD <- data.frame(colData(sce))
# pD <- left_join(pD,pcs.all[,c("Condition","ptime")])
# colnames(pD)[colnames(pD)=="ptime"] <- "ptime"
# pD <- pD[pD$MajorGroups %in% "Fibroblast",]
pD <- droplevels(pD)
sumry <- group_by(pD, ptimeBin, CellTypes) %>%
    summarize(n=n()) %>%
    mutate(frac=n/sum(n)) %>%
    ungroup() # %>%
#     mutate(cumfrac=cumsum(frac)) %>%
#     ungroup()

# ggplot(sumry, aes(x=ptime, y=frac)) +
#     geom_point() +
#     facet_wrap(~CellTypes)

comb <- table(sumry$CellTypes,sumry$ptimeBin) 
missing.comb <- which(comb==0, arr.ind=TRUE)
if (nrow(missing.comb)>0) {
    add.missing <- data.frame("CellTypes"=rownames(missing.comb),
		      "ptimeBin"=colnames(comb)[missing.comb[,"col"]],
		      "frac"=0,
		      "n"=0)
    sumry <- rbind(sumry,add.missing)
}
sumry$ptimeBin <- as.numeric(sumry$ptimeBin)

sumry$Colors <- plyr::mapvalues(as.character(sumry$CellTypes),
				as.character(unique(pD$CellTypes)),
				as.character(unique(pD$Colors)))

combi <- unique(paste0(substr(pD$MajorGroups,1,3),"-",as.character(pD$CellTypes)))
names(combi) <- unlist(lapply(strsplit(combi,"-"),function(x) x[[2]]))
sumry$CellTypes <- plyr::mapvalues(as.character(sumry$CellTypes),
				   as.character(names(combi)),
				   as.character(combi))

sumry <- arrange(sumry, ptimeBin, CellTypes) %>%
    group_by(ptimeBin) %>%
    mutate(cumfrac=cumsum(frac)) %>%
    ungroup()

barplt <- ggplot(sumry, aes(x=factor(ptimeBin,levels=c(5,4,3,2,1)), y=frac, fill=CellTypes)) +
    geom_bar(stat="identity",color="black") +
    scale_fill_manual(values=unique(sumry$Colors)) +
    xlab("Tumortime Bin") +
    ylab("Fraction of Cells") +
    coord_flip()

p4 <- ggplot(sumry, aes(x=as.numeric(ptimeBin), ymin=cumfrac-frac, ymax=cumfrac,
		  fill=CellTypes)) +
    geom_ribbon() +
    scale_fill_manual(values=unique(sumry$Colors)) +
    xlab("Tumor Development") +
    ylab("Fraction of Cells") +
    theme(legend.position="bottom")
```


# Figure
This will obviously still need the schematic
```{r, message=FALSE, warning=FALSE, fig.width=7, fig.height=12}

barplt <- barplt + theme(legend.position="none")
ab <- plot_grid(p1_col,p.ld.2,labels="auto",ncol=2)
abc <- plot_grid(ab,p0,ncol=1,labels=c("","c"),rel_heights=c(1,1.5))
d <- plot_grid(barplt,NULL,ncol=2)
plot_grid(abc,d,rel_heights=c(2.5,1),ncol=1)
```

#Saving individual plots

```{r, message=FALSE,warning=FALSE}
ages <- group_by(pD, Age, Condition) %>%
    summarize()
ptime <- left_join(ptime,ages)
ptime$Age <- as.numeric(ptime$Age)
ptime$ptimeBin <- as.factor(ptime$ptimeBin)

ageplot <- ggplot(ptime, aes(x=ptimeBin, y=Age)) +
    stat_summary(fun.data="mean_sdl", fun.args = list(mult=1), 
	    geom="errorbar", color="black", width=0.2) +
    stat_summary(fun.y=mean, geom="point", color="black") 
ageplot
```

```{r, message=FALSE, warning=FALSE}
outdr <- "../../data/figures/Figure1/"
ggsave(paste0(outdr,"BarplotTumor.svg"),barplt)
ggsave(paste0(outdr,"PCA.svg"),p1_col)
ggsave(paste0(outdr,"AgePlot.svg"),ageplot)
ggsave(paste0(outdr,"GOTerms.svg"),p.ld.2)
```

```{r, message=FALSE, warning=FALSE,fig.width=7,fig.height=7}
dpi <- 1200

ggplot(pD, aes(x=UMAP1, y=UMAP2, color=CellTypes)) +
    geom_point_rast(size=.3, dpi=dpi) +
    scale_color_manual(values=levels(pD$Color)) +
    theme_void() +
    theme(legend.position="none")

ggsave(paste0(outdr,"CellTypesUMAP.png"),dpi=dpi)

# With Annotation
X <- model.matrix(~0+pD$CellTypes)
colnames(X) <- levels(pD$CellTypes)
X <- t(t(X)/colSums(X))
celltyp_ctrs <- data.frame(crossprod(X, as.matrix(pD[,c("UMAP1","UMAP2")])))
celltyp_ctrs$CellType <- rownames(celltyp_ctrs)

ggplot(pD, aes(x=UMAP1, y=UMAP2, color=CellTypes)) +
    geom_point_rast(size=.3, dpi=dpi) +
    scale_color_manual(values=levels(pD$Color)) +
    geom_text_repel(data=celltyp_ctrs,aes(label=CellType,color=NULL)) +
    theme(legend.position="none")
ggsave(paste0(outdr,"CellTypesUMAP_WithLabels.png"),dpi=dpi)

```

# Split UMAPs into individual files
```{r, message=FALSE, warning=FALSE,fig.width=7,fig.height=7}
pltUMAP <- function(pD.sub,pD) {
    p <- ggplot(pD, aes(x=UMAP1, y=UMAP2)) +
	geom_point_rast(color="grey80",size=.3,dpi=dpi) +
	geom_point_rast(data=pD.sub, aes(color=Groups), size=.3,dpi=900) +
	scale_color_manual(values=levels(pD.sub$GroupColors)) +
	theme_void() +
	theme(legend.position="none")
    return(p)
}


pD$GroupColors <- plyr::mapvalues(pD$Groups,
				levels(pD$Groups),
				c("#00BFFF","#37A055","#3787C0","#F78757","#FF4E03",
				  "#737373","#800080"))
pDWt <- data.frame(colData(sce.full[,sce.full$Condition %in% c("CBLA1","CBLT","CTRL")]))
pDWt$GroupColors <- plyr::mapvalues(pDWt$Groups,
				levels(pDWt$Groups),
				c("#00BFFF","#37A055","#3787C0","#F78757","#FF4E03",
				  "#737373","#800080"))

set.seed(42)
# Select each pD individually and downsample
pD1 <- pD[pD$ptimeBin==1,]
pD1 <- pD1[sample(nrow(pD1),12000),]

pD2 <- pD[pD$ptimeBin==2,]
pD2 <- pD2[sample(nrow(pD2),12000),]

pD3 <- pD[pD$ptimeBin==3,]
pD3 <- pD3[sample(nrow(pD3),12000),]

pD4 <- pD[pD$ptimeBin==4,]
pD4 <- pD4[sample(nrow(pD4),12000),]

pD5 <- pD[pD$ptimeBin==5,]

pDWtO <- pDWt[pDWt$Condition %in% c("CBLA1","CBLT"),]
pDWtO <- pDWtO[sample(nrow(pDWtO),12000),]
pDWtY <- pDWt[pDWt$Condition %in% c("CTRL"),]
pDWtY <- pDWtY[sample(nrow(pDWtY),12000),]


ggsave(paste0(outdr,"UMAP_AllTumorigenesis.png"),pltUMAP(pD,pD),dpi=dpi)
ggsave(paste0(outdr,"UMAP_Bin_1.png"),pltUMAP(pD1,pD),dpi=dpi)
ggsave(paste0(outdr,"UMAP_Bin_2.png"),pltUMAP(pD2,pD),dpi=dpi)
ggsave(paste0(outdr,"UMAP_Bin_3.png"),pltUMAP(pD3,pD),dpi=dpi)
ggsave(paste0(outdr,"UMAP_Bin_4.png"),pltUMAP(pD4,pD),dpi=dpi)
ggsave(paste0(outdr,"UMAP_Bin_5.png"),pltUMAP(pD5,pD),dpi=dpi)
ggsave(paste0(outdr,"UMAP_Bin_WT_Young.png"),pltUMAP(pDWtY,pD),dpi=dpi)
ggsave(paste0(outdr,"UMAP_Bin_WT_Old.png"),pltUMAP(pDWtO,pD),dpi=dpi)

plot_grid(pltUMAP(pD1,pD) + ggtitle("1"),
	  pltUMAP(pD2,pD)+ ggtitle("2"),
	  pltUMAP(pD3,pD)+ ggtitle("3"),
	  pltUMAP(pD4,pD)+ ggtitle("4"),
	  pltUMAP(pD5,pD)+ ggtitle("5"),
	  pltUMAP(pDWtY,pD)+ ggtitle("WtY"),
	  pltUMAP(pDWtO,pD)+ ggtitle("WtO"))
guides(colour=guide_legend(title.position="top", override.aes = list(size=3))) +
    theme(legend.position="bottom",
	  legend.direction="horizontal",
	  strip.text=element_text(size=14))
```
