---
title: "RANK signaling"
author: "Karsten Bach"
date: '`r Sys.Date()`'
output:
  html_notebook:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: united
    highlight: tango
    code_folding: hide
---
***

# Introduciton

__RANKL(Tnfsf11)-RANK(Tnfrsf11a)__ signalling is a critical downstream pathway of progesterone.
Upon exposure to progesterone, first Pgr+ cells expand followed by a more substantial wave of proliferation of Pgr- cells. The second wave is thought be a result of the release of paracrine factors from the Pgr+ cells, mainly RANKL.

It has later been recognized that RANKL is deregulated in BRCA1 carriers. First by Visvader: 

> We identified two subsets of luminal progenitors (RANK + and RANK − ) in histologically normal tissue of BRCA1-mutation carriers and showed that RANK + cells are highly proliferative, have grossly aberrant DNA repair and bear a molecular signature similar to that of basal-like breast cancer.
> These data suggest that RANK + and not RANK − progenitors are a key target population in these women.
> Inhibition of RANKL signaling by treatment with denosumab in three-dimensional breast organoids derived
> from pre-neoplastic BRCA1 mut/+ tissue attenuated progesterone-induced proliferation.
Notably, proliferation was markedly reduced in breast biopsies from BRCA1-mutation carriers who were treated with denosumab.
> Furthermore, inhibition of RANKL in a Brca1-deficient mouse model substantially curtailed mammary tumorigenesis.

And also by Josef Penninger:

> Here, using two different mouse models, we show that genetic inactivation of the key osteoclast differentiation factor RANK in the mammary epithelium markedly delayed onset, reduced incidence, and attenuated progression of Brca1;p53 mutation-driven mammary cancer. Long-term pharmacological inhibition of the RANK ligand RANKL in mice abolished the occurrence of Brca1 mutation-driven pre-neoplastic lesions. Mechanistically, genetic inactivation of Rank or RANKL/RANK blockade impaired proliferation and expansion of both murine Brca1;p53 mutant mammary stem cells and mammary progenitors from human BRCA1 mutation carriers. In addition, genome variations within the RANK locus were significantly associated with risk of developing breast cancer in women with BRCA1 mutations. Thus, RANKL/ RANK control progenitor cell expansion and tumorigenesis in inherited breast cancer.

This has led to (several?) clinical trials using denosumab (neutralizing RANKL mAb) as chemopreventative agent in BRCA1 carriers.
I couldn't find many more follow up studies that addressed where the RANKL was coming from apart from [one study](https://pubmed.ncbi.nlm.nih.gov/28388533/?from_linkname=pubmed_pubmed_citedin&from_from_uid=27322743&from_page=7&from_pos=10) in Oncotarget which suggest it was cell-autonomous. 
Hence from my understanding the current standard of knowledge is:

- RANK is expressed on the cell-of-origin of BRCA1 associated BRCA
- RANK+ cells are highly proliferative and show DNA damage
- Denosumab reduces proliferation in BRCA1-mutation carriers
- Denosumab in Brca1-deficient mouse models delays tumorigenesis 
- Deleting RANK delays tumorigenesis
- In vitro data suggests that this is cell autonomous


```{r, message=FALSE,warning=FALSE}
library(slingshot)
library(scran)
library(scater)
library(ggplot2)
library(cowplot)
library(Matrix)
library(umap)
library(dplyr)
library(viridis)
source("../functions.R")
theme_set(theme_cowplot())

sce.full <- readRDS("../../data/Integrated//Robjects/SCE_combined_final.rds")
sce.full$SubCellTypes <- factor(sce.full$CellTypes)
sce.full$CellTypes <- factor(sce.full$CellTypesFinal)
sce.full$Colors <- factor(sce.full$Colors)
sce.full$Batch <- as.character(sce.full$Batch)
sce.full$Groups <- factor(sce.full$Groups)
fkingp53_andtm <- c("WKBR76.5c","WKBR76.5f","WKBR61.4dTM","WKBR75.4bTM")
sce.full <- sce.full[,sce.full$MajorGroups=="Epithelial" & !(sce.full$SampleID %in% fkingp53_andtm)]
sce <- sce.full[,sce.full$Experiment=="Tumorigenesis"]

pD <- droplevels(data.frame(colData(sce)))

rD <- rowData(readRDS("../../data/Tumorigenesis/Robjects/SCE.rds"))
rmgenes <- rownames(rD)[!rD$KeepForHvg]
rmgenes <- rownames(sce)[rownames(sce) %in% rmgenes]

dirtyPlot <- function(fplot){
    sub.gse <- quickGSE(fplot$Gene,fplot$fFDR)
    sub.gse$Fisher[sub.gse$Fisher=="< 1e-30"] <- "1e-30"
    sub.gse$Fisher <- -log10(as.numeric(sub.gse$Fisher))
    sub.gse %>%
	mutate(hitsPerc=(Significant/200)*100) %>% 
			    ggplot(aes(y=hitsPerc, 
				       x=Term, 
				       fill=Fisher
				       )) +
				geom_col() +
				expand_limits(x=0) +
				labs(y="Hits (%)", x="GO term", colour="p value", size="Count") +
				geom_text(aes(label=Term),y=1,
					  hjust=0,
					  color="grey30") +
				ylim(c(0,50)) +
				scale_fill_viridis(begin=1,end=0) +
				theme(axis.text.y=element_blank(),
				      axis.ticks.y=element_blank()) +
				coord_flip() 
}
outdr <- "../../data/figures/Figure2/"
```

# Epithelial Data
Just as a reminder this the epithelial data that we have.
```{r, fig.width=10, fig.height=5, warning=FALSE, message=FALSE}
p0 <- ggplot(pD, aes(x=UMAP1, y=UMAP2)) +
    geom_point(data=pD[,c("UMAP1","UMAP2")], size=.1,color="grey80") +
    geom_point(size=.1,aes(color=CellTypesFinal)) +
    ggtitle("Tumorigenesis") +
    scale_color_manual(values=levels(pD$Colors)) +
    facet_wrap(~Condition,nrow=2) +
    theme(legend.position="none") +
    theme_void()

sce.p <- sce.full[,sce.full$Experiment=="Pregnancy"]
pD.p <- data.frame(colData(sce.p))

p1 <- ggplot(pD.p, aes(x=UMAP1, y=UMAP2)) +
    geom_point(data=pD.p[,c("UMAP1","UMAP2")], size=.1,color="grey80") +
    geom_point(size=.1,aes(color=CellTypesFinal)) +
    ggtitle("Pregnancy") +
    scale_color_manual(values=levels(pD.p$Colors)) +
    facet_wrap(~Condition,nrow=2) +
    theme(legend.position="none") +
    theme_void()

p1
p0
```

# Points to make in the paper

## RANKL is de-regulated in BRCA1 animals
We can definitely show that RANKL expression is de-regulated in the BRCA1 mice and that it is expressed by the hormone-sensing cells.
With de-regulated I mean higher than in NP WT.

```{r, fig.width=7, fig.height=7, warning=FALSE, message=FALSE}
library(viridis)
library(schex)
sce <- make_hexbin(sce, nbins = 50, 
		   dimension_reduction = "UMAP", use_dims=c(1,2))
sce.p <- make_hexbin(sce.p, nbins = 50, 
		   dimension_reduction = "UMAP", use_dims=c(1,2))

p0 <- plot_hexbin_gene(sce, gene="Tnfsf11", type="logcounts",
		 action="mean", xlab="UMAP1", ylab="UMAP2") +
ggtitle("Rankl in T")
p1 <- plot_hexbin_gene(sce.p, gene="Tnfsf11", type="logcounts",
		 action="mean", xlab="UMAP1", ylab="UMAP2") +
ggtitle("Rankl in P")
p2 <- plot_hexbin_gene(sce, gene="Tnfrsf11a", type="logcounts",
		 action="mean", xlab="UMAP1", ylab="UMAP2") +
	ggtitle("Rank in T")
p3 <- plot_hexbin_gene(sce.p, gene="Tnfrsf11a", type="logcounts",
		 action="mean", xlab="UMAP1", ylab="UMAP2") +
	ggtitle("Rank in P")
plot_grid(p0,p1,p2,p3)
```

## Csn2 expression precedes Rankl expression in BRCA1 animals
The plot shows the expression of Csn2 (in Lps/Av) and Rankl in hormone sensing cells.
The fact that we see substantial Csn2 expression in BRCA1 animals before we see levels above WT NP Rankl levels in Hs cells suggests that Rankl is a secondary event.
```{r, fig.width=12, fig.height=4, warning=FALSE, message=FALSE}
sce.full$CT <- as.character(sce.full$CellTypesFinal)
sce.full$CT[sce.full$CT %in% c("Lp","Avd")] <- "LpAv"
sce.full$CT <- as.factor(sce.full$CT)

### With multiple genes
genesAv <- c("Csn2","Csn1s1","Csn1s2a","Wap","Glycam1")
genesHs <- c("Igf2","Tnfsf11","Wnt4","Areg")
genes <- c(genesAv[c(1,4)],genesHs[c(1,2)])
out <-  data.frame()
for (gen in genes) {
    ct <- ifelse(gen %in% genesAv,"LpAv","Hs")
    tmp <- data.frame(colData(sce.full[,sce.full$CT %in% ct]))
    tmp$Expr <- logcounts(sce.full[,sce.full$CT %in% ct])[gen,]
    tmp$GeneCT <- paste0(ct,"-",gen)
    out <- rbind(out,tmp)
}
ggplot(out, aes(x=Sample,y=Expr))+
    geom_violin() +
    geom_jitter(size=.1) +
    facet_grid(GeneCT~Condition,scales="free") +
    theme(strip.text.x = element_text(size=12, face="bold") ,
          strip.text.y = element_text(size=12, face="bold"),
          strip.background = element_rect(colour="black", fill="white"),
	  axis.text.x = element_blank())
ggsave(paste0(outdr,"MultipleGenes-Violin-Plot.pdf"),width=8,height=6)
```

### The same as bubbleplot?
```{r, fig.width=12, fig.height=4, warning=FALSE, message=FALSE}
genesAv <- c("Csn2","Csn1s1","Csn1s2a","Wap","Glycam1")
genesHs <- c("Igf2","Tnfsf11")

getMtx <- function(m,markers,grps) {
    require(reshape2)
    mtx <- data.frame(numeric(length(markers)))
    colnames(mtx) <- levels(factor(grps))[1]
    mtx.freq <- mtx
    for (cond in levels(factor(grps))) {
	expr <- rowMeans(m[markers,grps==cond])
	expr.freq <- rowMeans(m[markers,grps==cond]>0)
	colname <- cond
	mtx[,colname] <- expr
	mtx.freq[,colname] <- expr.freq
    }
    mtx <- as.matrix(mtx)
    mtx.freq <- as.matrix(mtx.freq)
    rownames(mtx.freq) <- rownames(mtx) <- markers
    mtx <- mtx/rowMax(mtx)
    mtx.long <- melt(mtx,value.name="Mean")
    mtx.freq <- melt(mtx.freq,value.name="Frequency")
    mtx.long$Frequency <- mtx.freq$Frequency * 100
    return(mtx.long)
}

out.Av <- getMtx(assay(sce.full[,sce.full$CT=="LpAv"],"logcounts"), genesAv,
       colData(sce.full[,sce.full$CT=="LpAv"])$SampleID)

out.Hs <- getMtx(assay(sce.full[,sce.full$CT=="Hs"],"logcounts"), genesHs,
       colData(sce.full[,sce.full$CT=="Hs"])$SampleID)

out.long <- rbind(out.Av, out.Hs)
colnames(out.long) <- c("Gene","SampleID","Mean","Frequency")

add <- data.frame(colData(sce.full))[,c("SampleID","Condition")] %>%
    group_by(SampleID,Condition) %>%
    summarize() %>%
    ungroup()

add2 <- data.frame("Gene"=c(genesAv,genesHs),
		   "CellType"=c(rep("LpAv",length(genesAv)),
				rep("Hs",length(genesHs))))

out.long <- dplyr::left_join(out.long,add)
out.long <- dplyr::left_join(out.long,add2)

ggplot(out.long[out.long$Condition %in% c("WTYoung","4.5dG","9.5dG","14.5dG"),], aes(x=SampleID, y=Gene, color=Mean, size=Frequency)) +
    geom_point() +
    scale_color_gradient2(high="red",limits=c(0,1)) + ##,mid="orange") +
   # scale_color_distiller(palette="Spectral") +
    scale_size(range=c(0,3)) +
    theme(panel.grid.major=element_line(colour="grey80",size=0.1,linetype="dashed"),
	  axis.text=element_text(size=14),
	  legend.position="bottom",
	  legend.direction="horizontal",
	  strip.text.x = element_text(size=12, face="bold") ,
	  strip.text.y = element_text(size=12, face="bold"),
	  strip.background = element_rect(colour="black", fill="white"),
	  axis.text.x=element_blank(),
	  axis.ticks.x=element_blank()) +
    facet_grid(CellType~Condition,scales="free",space="free_y") +
    xlab("") +
    ylab("")
ggsave(paste0(outdr,"MultipleGenes-Bubble-Plot_Pregnancy.pdf"),width=6.77,height=3.27)


ggplot(out.long[out.long$Condition %in% c("WTOld","1","2","3","4"),], aes(x=SampleID, y=Gene, color=Mean, size=Frequency)) +
    geom_point() +
    scale_color_gradient2(high="red",limits=c(0,1)) + ##,mid="orange") +
   # scale_color_distiller(palette="Spectral") +
    scale_size(range=c(0,3)) +
    theme(panel.grid.major=element_line(colour="grey80",size=0.1,linetype="dashed"),
	  axis.text=element_text(size=14),
	  legend.position="bottom",
	  legend.direction="horizontal",
	  strip.text.x = element_text(size=12, face="bold") ,
	  strip.text.y = element_text(size=12, face="bold"),
	  strip.background = element_rect(colour="black", fill="white"),
	  axis.text.x=element_blank(),
	  axis.ticks.x=element_blank()) +
    facet_grid(CellType~Condition,scales="free",space="free_y") +
    xlab("") +
    ylab("")
ggsave(paste0(outdr,"MultipleGenes-Bubble-Plot_Tumorigenesis.pdf"),width=6.77,height=3.27)
```

### The same as line plot?
```{r, fig.width=12, fig.height=4, warning=FALSE, message=FALSE}
genesAv <- c("Csn2","Wap")
genesHs <- c("Igf2","Tnfsf11")

getMtx <- function(m,markers,grps) {
    require(reshape2)
    mtx <- data.frame(numeric(length(markers)))
    colnames(mtx) <- levels(factor(grps))[1]
    mtx.freq <- mtx
    for (cond in levels(factor(grps))) {
	expr <- rowMeans(m[markers,grps==cond])
	expr.freq <- rowMeans(m[markers,grps==cond]>0)
	colname <- cond
	mtx[,colname] <- expr
	mtx.freq[,colname] <- expr.freq
    }
    mtx <- as.matrix(mtx)
    mtx.freq <- as.matrix(mtx.freq)
    rownames(mtx.freq) <- rownames(mtx) <- markers
    mtx <- mtx/rowMax(mtx)
    mtx.long <- melt(mtx,value.name="Mean")
    mtx.freq <- melt(mtx.freq,value.name="Frequency")
    mtx.long$Frequency <- mtx.freq$Frequency * 100
    return(mtx.long)
}

out.Av <- getMtx(assay(sce.full[,sce.full$CT=="LpAv"],"logcounts"), genesAv,
       colData(sce.full[,sce.full$CT=="LpAv"])$SampleID)

out.Hs <- getMtx(assay(sce.full[,sce.full$CT=="Hs"],"logcounts"), genesHs,
       colData(sce.full[,sce.full$CT=="Hs"])$SampleID)

out.long <- rbind(out.Av, out.Hs)
colnames(out.long) <- c("Gene","SampleID","Mean","Frequency")

add <- data.frame(colData(sce.full))[,c("SampleID","Condition","ptime")] %>%
    group_by(SampleID,Condition,ptime) %>%
    summarize() %>%
    ungroup()

add2 <- data.frame("Gene"=c(genesAv,genesHs),
		   "CellType"=c(rep("LpAv",length(genesAv)),
				rep("Hs",length(genesHs))))

out.long <- dplyr::left_join(out.long,add)
out.long <- dplyr::left_join(out.long,add2)

out.long$Experiment <- ifelse(out.long$Condition %in% c("WTYoung","4.5dG","9.5dG","14.5dG"),
			      "Pregnancy",
			      "Tumourigenesis")
out.long$Xaxis <- ifelse(out.long$Experiment=="Pregnancy",as.numeric(out.long$Condition)-1,
			     as.numeric(out.long$ptime))
out.long$Xaxis[out.long$Xaxis==2] <- 4.5
out.long$Xaxis[out.long$Xaxis==3] <- 9.5
out.long$Xaxis[out.long$Xaxis==4] <- 14.5
out.long$Xaxis[is.na(out.long$Xaxis)] <- 0

ggplot(out.long, aes(x=Xaxis, y=Mean,  group=Gene,color=Gene)) +
    geom_point() +
    geom_smooth(se=FALSE,aes(lty=CellType),lwd=2) +
   # scale_color_distiller(palette="Spectral") +
    theme_pub() +
    theme(panel.grid.major=element_line(colour="grey80",size=0.1,linetype="dashed"),
	  axis.text=element_text(size=14),
	  legend.position="bottom",
	  legend.direction="horizontal",
	  strip.text.x = element_text(size=12, face="bold") ,
	  strip.text.y = element_text(size=12, face="bold"),
	  strip.background = element_rect(colour="black", fill="white"),
	  )+
    facet_grid(~Experiment,scales="free",space="free_y") +
    scale_color_manual(values=c("#6BAED6","#FEE0D2","#FC9272","#08519C")) +
    xlab("") +
    ylab("")
ggsave(paste0(outdr,"CsnRanklRegressionPlot.pdf"),width=9.61,height=5.87)
```

## Alveogenesis is not hormonally driven
Well I would like to make this point, I think the RANKL already does it but it would be nice to have some sort of Pr/Prl signature to show there is no hormonal response in the Hs cells.
The one I have right now is from a cancer cell line and is clearly rubbish. 
Struggling to find anything useful.
```{r, fig.width=12, fig.height=4, warning=FALSE, message=FALSE,eval=FALSE}
# From a cancer cell line (MigSigDB)
respGenes <- c("Abcc5","Ar","Atf3","Bcar1","Bcas2","Bccip","Bnip3","Ccn1","Ccnd1",
	       "Cdkn1c","Cebpd","Dnah1","Dusp1","Eci2","Egf","Gadd45a","Gm2a",
	       "Grb2","Hdac9","Hmgb3","Il6st","Jun","Kat2b","Map3k3","Map4k5",
	       "Nme2","Polr2k","Rasl10b","Sap30","Sos1","Tgfa","Tp53bp2","Uchl5")
respGenes <- respGenes[respGenes %in% rownames(sce.full)]

# From an organoid paper
respGenes2 <- c("Rgs2","Defb1","Saa1","Saa3","Saa2","Plac8","Tnfsf11","Fgg",
		"Agtr1","Sult1a1","Rasd1","Cited1","Slc25a29","Spdef",
		"Sftpd","Gpx3","Tnxb","Apof","Pla2g5","Vwf",
		"Ly6c","Ly6a","Tgm1","Ly6f","Ly64","Ccdc92",
		"Wnt4","Cdo1","Ifi44")
respGenes2 <- respGenes2[respGenes2 %in% rownames(sce.full)]

# From GO
library(org.Mm.eg.db)
rp.genes <- select(org.Mm.eg.db, keys="GO:0032570", keytype="GOALL", column="ENSEMBL")

rp.genes <- data.frame(rD)[rD$ID %in% rp.genes$ENSEMBL,"uniq"]
rp.genes <- rp.genes[rp.genes %in% rownames(sce.full)]

fplot$PrResponse <- colSums(assays(sce.full[,sce.full$CT %in% c("LpAv","Hs")])[["reconstructed"]][rp.genes,])

ggplot(fplot, aes(x=Sample,y=PrResponse))+
    geom_violin(draw_quantiles=0.5) +
    #     geom_jitter(size=.3) +
    facet_grid(CT~Condition,scales="free") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Which receptor binding genes are up-regulated
```{r, fig.width=10, fig.height=10, warning=FALSE, message=FALSE}
library(org.Mm.eg.db)
markers <- findMarkers(sce,factor(sce$CellTypesFinal),
		       direction="up",
		       pval.type="all",
		       subset.row=!rownames(sce) %in% rmgenes)
av.mrk <- data.frame(markers[["Avd"]])
upgenes <- rownames(av.mrk)[av.mrk$FDR<0.01]

sec.genes <- select(org.Mm.eg.db, keys="GO:0005102", keytype="GOALL", column="ENSEMBL")
sec.genes <- data.frame(rD)[rD$ID %in% sec.genes$ENSEMBL,"uniq"]
sec.genes <- sec.genes[sec.genes %in% rownames(sce)]

feats <- upgenes[upgenes %in% sec.genes]
sce$CellTypesFinal <- droplevels(sce$CellTypesFinal)
print(feats)
plotExpression(sce, x="CellTypesFinal", feats) + ggtitle("Genes Up")
```

### Receptors of some of the genes
I looked at the following l-r pairs:
(There are also paper showing that Ccl20 up-regulates RANKL)

- Tnfs15-Tnfrs25
- Saa1 - Fpr2
- Rspo1 - Lgr4-6
- Ccl20 - Ccr6, Cxcr3
```{r, fig.width=10, fig.height=10, warning=FALSE, message=FALSE}
feats <- c("Tnfrsf25","Fpr2","Lgr4","Lgr5","Lgr6","Ccr6","Cxcr3")
plotExpression(sce, x="CellTypesFinal", feats)
```

## Cytokines/Morphogens known to up-regulated RANKL
This is from a review from [Penninger](https://nyaspubs.onlinelibrary.wiley.com/doi/10.1196/annals.1443.016)
```{r, warning=FALSE, message=FALSE}
whatabout <- c("Tnf","Il1b","Il6","Il11","Il17a","Bmp2",
	       "Lif","Igf1")
plotExpression(sce, x="CellTypesFinal", whatabout) + ggtitle("Cytokines/Morphogenes known to up-regulated RANKL")
feats <- c("Lifr")
plotExpression(sce, x="CellTypesFinal", feats)

p0 <- plot_hexbin_gene(sce, gene="Lif", type="logcounts",
		 action="mean", xlab="UMAP1", ylab="UMAP2") +
ggtitle("Lif in T")
p1 <- plot_hexbin_gene(sce, gene="Lifr", type="logcounts",
		 action="mean", xlab="UMAP1", ylab="UMAP2") +
ggtitle("Lifr in P")
```


### DE Analysis of early LPs
```{r, fig.width=7, fig.height=7, warning=FALSE, message=FALSE}
sce.sub <- sce.full[,sce.full$Condition %in% c("1","WTOld") & sce.full$CellTypesFinal=="Lp"]
sce.sub$Condition <- droplevels(sce.sub$Condition)
table(sce.sub$Sample)
smrzd <- aggregateAcrossCells(sce.sub,
			      id=colData(sce.sub)[,c("Sample")])
library(edgeR)
y <- DGEList(counts=counts(smrzd),
	     samples=colData(smrzd))
levels(y$samples$Condition) <- c("WT","WKBR")
keep <- filterByExpr(y, group=y$samples$Condition)
y <- y[keep,]
# Norm
y <- calcNormFactors(y)
mdl <- model.matrix(~ Condition,y$samples)
y <- estimateDisp(y, mdl)

fit <- glmQLFit(y, mdl, robust=TRUE)
res <- glmQLFTest(fit, coef=ncol(mdl))
summary(decideTests(res))

top <- topTags(res,n=nrow(y))$table
top$Gene <- rownames(top)
library(ggrepel)
ggplot(top, aes(x=logFC, y=-log10(FDR))) +
    geom_point(size=.5,color="grey80") +
    geom_point(data=top[top$FDR<0.1,], size=.9)+
    geom_label_repel(data=top[top$FDR<0.1,][1:30,], aes(label=Gene)) +
    theme(text=element_text(size=12)) +
    xlab("logFC") +
    ylab("-log10(FDR)")
ggsave(paste0(outdr,"Volcano_Bin1-WT_LPonly.pdf"))

plotExpression(sce.full[,sce.full$CellTypesFinal=="Lp"],x="Condition","Cdkn2a")
top$fFDR <- ifelse(top$FDR<0.1 & top$logFC>0,0.001,1)
dirtyPlot(top) + ggtitle("Up in Bin1")
top$fFDR <- ifelse(top$FDR<0.1 & top$logFC<0,0.001,1)
dirtyPlot(top) + ggtitle("Up in WT")
```

### Changes in Lps over time
```{r, fig.width=10, fig.height=10, warning=FALSE, message=FALSE}
sce.sub <- sce.full[,sce.full$Condition %in% c("1","2","3","4") & sce.full$CellTypesFinal=="Lp"]
sce.sub$Condition <- droplevels(sce.sub$Condition)
table(sce.sub$Sample)
smrzd <- aggregateAcrossCells(sce.sub,
			      id=colData(sce.sub)[,c("Sample")])
library(edgeR)
y <- DGEList(counts=counts(smrzd),
	     samples=colData(smrzd))
y$samples$Condition <- as.numeric(y$samples$Condition)
keep <- filterByExpr(y, group=y$samples$Condition)
y <- y[keep,]
# Norm
y <- calcNormFactors(y)
mdl <- model.matrix(~ Condition,y$samples)
y <- estimateDisp(y, mdl)

fit <- glmQLFit(y, mdl, robust=TRUE)
res <- glmQLFTest(fit, coef=ncol(mdl))
summary(decideTests(res))

top <- topTags(res,n=nrow(y))$table
top$Gene <- rownames(top)
library(ggrepel)
ggplot(top, aes(x=logFC, y=-log10(FDR))) +
    geom_point(size=.5,color="grey80") +
    geom_point(data=top[top$FDR<0.05,], size=.9)+
    geom_label_repel(data=top[1:40,], aes(label=Gene)) +
    geom_hline(yintercept=2) +
    ggtitle("DE w time")

top$fFDR <- ifelse(top$FDR<0.1 & top$logFC>0,0.001,1)
dirtyPlot(top) + ggtitle("Up with time")
top$fFDR <- ifelse(top$FDR<0.1 & top$logFC<0,0.001,1)
dirtyPlot(top) + ggtitle("Down with time")
```
