---
title: "Figure 1"
author: "Karsten Bach"
date: '`r Sys.Date()`'
output:
  html_notebook:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: united
    highlight: tango
    code_folding: hide
---
***

# Main messages

In this file I will produce all the figure files for F1.

__Experimental setup of Tumorigenesis experiment__

- There is a need to understand pre-cancerous changes
- Sampling animals from a TNBC model pre-tumor formation and performing scRNAseq

__Inferring biological time__

- Age does not capture disease progression
- Inference of tumor time and grouping into stages

__Compositional changes during tumor development__

- Facetted UMAP as Overview
- Possible Ribbonplot / Barplot
I will keep the code and the plots for all of this in this document.
Including some further discussions.

```{r, message=FALSE,warning=FALSE}
library(scran)
library(scater)
library(ggplot2)
library(cowplot)
library(dplyr)
library(edgeR)
library(Matrix)
library(ggrepel)
library(ggrastr)
library(viridis)
theme_set(theme_cowplot())

sce.full <- readRDS("../../data/Integrated//Robjects/SCE_combined_final.rds")

rD <- rowData(readRDS("../../data/Tumorigenesis/Robjects/SCE.rds"))
rmgenes <- rD$uniq[!rD$KeepForHvg]
outdr <- "../../data/figures/Supplementary/"
```


# UMAP
```{r, message=FALSE,warning=FALSE,fig.width=9,fig.height=6}
pD <- data.frame(colData(sce.full))
p <- ggplot(pD, aes(x=UMAP1, y=UMAP2, color=Batch)) +
    geom_point(size=.1) +
    theme_void() +
    theme(legend.position="bottom",
	  legend.direction="horizontal",
	  strip.text=element_text(size=14)) +
    scale_color_brewer(palette="Set1") +
    guides(colour = FALSE)
ggsave(paste0(outdr,"BatchCor.png"), dpi=400)
```

```{r, message=FALSE,warning=FALSE}
dec.full <- modelGeneVar(sce.full)
hvgs.full <- getTopHVGs(dec.full)
hvgs.full <- !hvgs.full %in% rmgenes

# PCA
set.seed(42)
sce <- denoisePCA(sce.full, technical=dec.full,
		       subset.row=hvgs.full,
		       BSPARAM=BiocSingular::IrlbaParam())
pcs <- reducedDims(sce)[["PCA"]]
library(umap)
ump <- umap(pcs, random_state=42)

pD$UMAP1Uncor <- ump$layout[,1]
pD$UMAP2Uncor <- ump$layout[,2]
p1 <- ggplot(pD, aes(x=UMAP1Uncor, y=UMAP2Uncor, color=Batch)) +
    geom_point(size=.1) +
    theme_void() +
    theme(legend.position="bottom",
	  legend.direction="horizontal",
	  strip.text=element_text(size=14)) +
    scale_color_brewer(palette="Set1") +
    guides(colour = guide_legend(override.aes = list(size=6)))
ggsave(paste0(outdr,"BatchUnCor.png"), dpi=400)
```
